---
title: "Cholera final runs"
author: "Jong-Hoon Kim"
date: "9/14/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load the package 

## Packages
```{r}
library(data.table)
library(devtools)
library(usethis)
library(tidyverse)
```

## Checklist
1. Download the data set and familiarize yourself with the systesm
2. Check if each function works for each data set provided
  a. Country names may differ between different data sets
  b. See if there are any missing values

## Steps
```{r setup, include=FALSE}
# load package
devtools::load_all()
countries <- target_countries_cholera
dis <- "cholera"

cntry <- countries[1]
pop <- setup_cohorts(country = cntry,
                     year = eval(parse(text = "2000:2100")),
                     population_data = population_data)

pop_list <- lapply(countries, function(x) 
  setup_cohorts(country = x, year = eval(parse(text = "2000:2100")),
                population_data = population_data))

# coverage_202110gavi-2_cholera-no-vaccination.csv
# coverage_202110gavi-2_cholera-campaign-default.csv
# vacc_cov_data <- 
#   fread("inst/extdata/coverage_202110gavi-2_cholera-campaign-default.csv")
# vacc_cov_data$country <- clean_country_names(vacc_cov_data$country)

vacc_cov_data <- vacc_cov_input_cholera
vc <- create_vaccine_coverage(country = cntry,
                        population = pop,
                        vaccine_coverage_data = vacc_cov_data)
vc_list <- list()
for (i in 1:length(countries)) {
  vc_list[[i]] <- create_vaccine_coverage(country = countries[[i]],
                        population = pop_list[[i]],
                        vaccine_coverage_data = vacc_cov_data)
}
nruns <- 30
params <- set_params_cholera(disease = dis,
                          nruns = nruns,
                          country = cntry,
                          parameter_data = parameter_data_cholera,
                          incidence_rate_data = overall_incid_rate_cholera,
                          case_fatality_ratio_data = cfr_data_cholera)

i <- 1

vacc_eff_by_year <- unlist(params[i, paste0("vacc_eff_year_", 0:6)])
vacc_protected <- 
  calculate_vaccine_protected(disease = dis,
                              country = cntry,
                              population = pop,
                              year = 2000:2100,
                              vaccine_coverage = vc,
                              exponential_decay = FALSE,
                              vaccine_efficacy_by_year = vacc_eff_by_year)

vacc_protect_list <- list()
for (i in 1:length(countries)) {
  cat("i =", i ,", country =", countries[[i]], "\n")
  vacc_protect_list[[i]] <-  calculate_vaccine_protected(disease = dis,
                              country = countries[[i]],
                              population = pop_list[[i]],
                              year = 2000:2100,
                              vaccine_coverage = vc_list[[i]],
                              exponential_decay = FALSE,
                              vaccine_efficacy_by_year = vacc_eff_by_year)
}

## population is now 
pop_unprotected <- pop - vacc_protected
sum(pop_unprotected < 0)

# sample an incidence rate
overall_incid_rate_cholera %>% 
  filter(country == cntry) -> ir

# lnorm_par <- calc_lognorm_pars(mean = ir_par$ir_sim, sd = ir_par$ir_sim_sd)
# 
# set.seed(1)
# overall_ir <- rlnorm(1, lnorm_par$mu, lnorm_par$sigma)  
  
# overall_ir_dat %>%
#   filter(country == cntry) %>%
#   select(min, max) %>% unlist() -> ir_val

set.seed(1)
overall_ir <- runif(1, min = 0, max = 1)
overall_ir_tr <- qlnorm(overall_ir , meanlog = ir$meanlog, sdlog = ir$sdlog)
  
incid_rate_adj <- data.frame(matrix(NA, ncol = 101))
names(incid_rate_adj) <- c(as.character(2000:2100))

wash_prop %>% filter(country == cntry) -> wash_prop_country

# risk_vars <- wash_risk_ratio$var
# [1] "at_least_basic_sanitation" "at_least_basic_water"     
# [3] "basic_hygiene"             "no_open_defecation"       
# [5] "no_surface_water"          "no_unimproved_sanitation" 
# [7] "no_unimproved_water"       "safely_managed_water" 
# 8 risk variables, but the first two risk variables were omitted because
# relevant variables from the case-control studies were not identified.

# remove columns with NAs

risk_vars <-
  names(wash_prop_country[,5:10])[!is.na(colSums(wash_prop_country[,5:10]))]

ir_adj <- 
  data.frame(matrix(NA, nrow = length(risk_vars), ncol = 101 + 1))
names(ir_adj) <- c("WASH_var", as.character(2000:2100))
ir_adj$WASH_var <- risk_vars

ref_year <- 2010
rv <- risk_vars[1]

wash_prop %>%
  filter(country == cntry, year == ref_year) %>%
  pull(rv) -> prop_low_risk_ref

wash_risk_ratio %>%
  filter(var == rv) %>%
  pull(risk_ratio) -> rr

ir_high <- overall_ir / (prop_low_risk_ref * rr + (1 - prop_low_risk_ref))
ir_low <- ir_high * rr

prop_low_risk <-
    wash_prop %>%
    filter(country == cntry) %>%
    pull(rv)

ir_adj[ir_adj$WASH_var == rv, 2:102] <-
        ir_high * (1 - prop_low_risk) +
        ir_low * prop_low_risk

for (rv in risk_vars) {
  ir_by_risk <- calc_incid_rate_by_risk(country = cntry,
                                        overall_ir = overall_ir,
                                        wash_risk_ratio = wash_risk_ratio,
                                        wash_prop = wash_prop,
                                        ref_year = 2010,
                                        risk_var = rv)
      # cat("country =", cntry, ", risk variable =", rv, "\n")
      # extract predicted proportion of low-risk people over 2000-2100
      wash_prop %>% 
        filter(country == cntry) %>%
        pull(rv) -> prop_low_risk

      ir_adj[ir_adj$WASH_var == rv, 2:102] <-
        ir_by_risk$ir_high * (1 - prop_low_risk) +
        ir_by_risk$ir_low * prop_low_risk
      # cat("ir =", ir_ref$ir_high, ir_ref$ir_low, "\n")
}

incid_rate_adj[1, ] <- apply(ir_adj[, 2:102], 2, mean, na.rm = TRUE)
for (r in 2:101) {
    incid_rate_adj[r, ] <- incid_rate_adj[1, ]
}

incid_rate_adj <- calc_incid_rate_risk_adj_chol(country = cntry,
                              overall_ir = overall_ir,
                              wash_risk_ratio = wash_risk_ratio,
                              wash_prop = wash_prop, 
                              ref_year = 2010)
  
incid_rate_adj_list <- list()
for (i in 1:length(countries)) {
  cat("i =", i ,", country =", countries[[i]], "\n")
  incid_rate_adj_list[[i]] <- calc_incid_rate_risk_adj_chol(country = countries[[i]],
                              overall_ir = overall_ir,
                              wash_risk_ratio = wash_risk_ratio,
                              wash_prop = wash_prop, 
                              ref_year = 2010)
}

## calculate_cases_tf can be used for cholera as well
# ir_adj <- incid_rate_adj_list[[1]]
ir_adj <- incid_rate_adj
cases <- calculate_cases(country = cntry,
                         population = pop_unprotected,
                         year = 2000:2100,
                         incidence_rate = ir_adj)

pop[1:10,1]
cases[1:10,1]
head(pop)
head(cases)

parameter

```

## Parameter sampling
```{r}
devtools::load_all()
disease <- "Cholera"
nruns <- 30
country <- target_countries_cholera[1]

params <- set_params_cholera(disease = disease,
                          nruns = nruns,
                          country = country,
                          parameter_data = parameter_data_cholera,
                          incidence_rate_data = overall_incid_rate_cholera,
                          case_fatality_ratio_data = cfr_data_cholera)

life_expectancy <- 
  get_life_expectancy(life_expectancy_data = life_expectancy,
                      country = country,
                      year = 2000:2100)

vacc_scenario <- "campaign"
i <- 1
parameter_sample <- params[i,]
vaccine_coverage_data <- vacc_cov_input_cholera

vacc_impact <- vaccine_impact_cholera(disease = disease,
                       country = country,
                       population_data = population_data,
                       wash_prop = wash_prop,
                       wash_risk_ratio = wash_risk_ratio,
                       vaccine_coverage_data = vacc_cov_input_cholera,
                       vacc_scenario = vacc_scenario,
                       year = 2000:2100,
                       life_expectancy = life_expectancy,
                       parameter_sample = parameter_sample)

```

## Check vaccine protected

```{r}
cntry <- "Bangladesh"
yr <- 2000:2100
vaccine_coverage_data <- vacc_cov_input_cholera
dis <- "Cholera"

pop <- setup_cohorts(country = cntry,
                     year = yr,
                     population_data = population_data)
  #
vacc_cov <-
      create_vaccine_coverage(country = cntry,
                              population = pop,
                              vaccine_coverage_data = vaccine_coverage_data)
colSums(vacc_cov)    

params_list <- readRDS("outputs/params_Cholera_20211220T130658.rds")
params <- params_list[[which(target_countries_cholera == "Bangladesh")]]
parameter_sample <- params[1,]

vacc_eff_by_year <- unlist(parameter_sample[paste0("vacc_eff_year_", 0:6)])
    #
vacc_protected <-
  calculate_vaccine_protected(disease = dis,
                              country = cntry,
                              population = pop,
                              vaccine_coverage = vacc_cov,
                              vaccine_efficacy = NULL,
                              vaccine_immunity_duration = NULL,
                              exponential_decay = FALSE,
                              vaccine_efficacy_by_year = vacc_eff_by_year)
colSums(vacc_protected)
vacc_recipients <-
  calculate_vaccine_protected(disease = dis,
                              country = cntry,
                              population = pop,
                              vaccine_coverage = vacc_cov,
                              vaccine_efficacy = 1.0,
                              vaccine_immunity_duration = 1000,
                              exponential_decay = TRUE,
                              vaccine_efficacy_by_year = NULL)


colSums(vacc_recipients)
pop_unprotected <- pop
# factor protection for vaccine recipients
pop_unprotected <- pop_unprotected - vacc_protected
# vaccine coverage used to switch on/off indirect effect
pop_unvacc <- pop - vacc_recipients
vacc_cov_year <-
  calc_vacc_cov_year(pop = pop, vacc_recipients = vacc_recipients)

  calc_indirect_vacc_protected_cholera(population = pop_unvacc,
                                       vacc_cov = vacc_cov_year)
colSums(vacc_recipients)
pop_unprotected <- pop_unprotected - indirect_vacc_protected
all.equal(colSums(pop)[1:21], colSums(pop_unprotected)[1:21])
```


### Create parameters list
```{r}
set.seed(1234)
params_list <- list()
nruns <- 200
dis <- "Cholera"
tstamp <- format(Sys.time(), "%Y%m%dT%H%M%S")
for (k in seq_along(target_countries_cholera)) {
  cntry <- target_countries_cholera[[k]]
  params_list[[k]] <-
    set_params_cholera(disease = dis,
                  nruns = nruns,
                  country = cntry,
                  parameter_data = parameter_data_cholera,
                  incidence_rate_data = overall_incid_rate_cholera,
                  case_fatality_ratio_data = cfr_data_cholera)
}
saveRDS(params_list, paste0("outputs/params_", dis, "_", tstamp, ".rds"))
```

## Simulation using simulate_cholera
```{r}
devtools::load_all()
params_list <- readRDS("outputs/params_Cholera_20211220T130658.rds")

# vacc_scenarios <- c("novacc", "campaign")
# vacc_scenarios <- c("novacc")
vacc_scenarios <- c("campaign")
tic <- Sys.time()
for (vs in vacc_scenarios){
  tstamp <- format(Sys.time(), "%Y%m%dT%H%M%S")
  dis <- "Cholera"
  sim <- simulate_cholera(country = target_countries_cholera,
                          params_list = params_list,
                          vacc_scenario = vs,
                          vacc_cov_data = vacc_cov_input_cholera,
                          life_expectancy_data = life_expectancy_data,
                          wash_prop = wash_prop,
                          wash_risk_ratio = wash_risk_ratio_cholera,
                          nruns = 200)
  
  message(paste0(Sys.time() - tic, " elapsed"))
  
  saveRDS(sim$stoch, paste0("outputs/stoch_output_list_", 
                 tstamp, "_", dis, "_", vs, ".rds"))
  saveRDS(sim$central, paste0("outputs/central_output_list_",
                 tstamp, "_", dis, "_", vs, ".rds"))
}
```


## Compare vaccine_impact vs. vaccine_impact_cholera
The difference comes from random numbers are different 
```{r}
# old version
cntry <- target_countries_cholera[1]
year_chr <- "2000:2100"
nruns <- 30
vacc_scenario <- "novacc"

cohort <- list()
cases <- list()
deaths <- list()
dalys <- list() 

life_exp <- 
  get_life_expectancy(life_expectancy_data = life_expectancy_data,
                      country = cntry,
                      year = eval(parse(text = year_chr)))

## country-specific parameters
params_list <- readRDS("outputs/params_Cholera_20211125T0734.rds")
params <- params_list[[1]] # country by index
for (i in 1:nruns) {
  res <- vaccine_impact(
      disease = dis,
      country = cntry,
      population = population_data,
      wash_prop = wash_prop,
      wash_risk_ratio = wash_risk_ratio_chol,
      vacc_scenario = vacc_scenario,
      vaccine_coverage = vacc_cov_input_cholera,
      year = eval(parse(text = year_chr)),
      life_expectancy = life_exp,
      incidence_rate = params$incidence_rate[i],
      disability_weight = params$disability_weight[i],
      case_fatality = params$case_fatality_ratio[i],
      illness_duration = 5/365, 
      vaccine_immunity_duration = NULL,
      exponential_decay = FALSE,
      vaccine_efficacy = params$ve_yr0[i],
      vaccine_efficacy_by_year = c(
        params$ve_yr0[i],
        params$ve_yr1[i],
        params$ve_yr2[i],
        params$ve_yr3[i],
        params$ve_yr4[i],
        params$ve_yr5[i],
        params$ve_yr6[i]))
  
  cohort[[i]] <- res$cohort_size
  cases[[i]] <- res$cases
  deaths[[i]] <- res$deaths
  dalys[[i]] <- res$dalys 
}
cases_sum <- Reduce('+', cases)
cases_sum_col <- colSums(cases_sum)
plot(2000:2100, cases_sum_col/nruns, type="l")
max(cases_sum_col/nruns)


# new version
cohort <- list()
cases <- list()
deaths <- list()
dalys <- list()
life_exp <-
  get_life_expectancy(life_expectancy_data = life_expectancy_data,
                      country = cntry,
                      year = eval(parse(text = year_chr)))

params_list <- readRDS("outputs/params_Cholera_20211220T130658.rds")
params <- params_list[[1]] # country by index
#params_list_old <- readRDS("outputs/params_Cholera_20211125T0734.rds")
# params_old <- params_list_old[[1]] # country by index
# params$incidence_rate[1:30] <- params_old$incidence_rate
for (i in 1:nruns) {
  res <- vaccine_impact_cholera(disease = dis,
                           country = cntry,
                           population_data = population_data,
                           wash_prop = wash_prop,
                           wash_risk_ratio = wash_risk_ratio_cholera,
                           vaccine_coverage_data = vacc_cov_input_cholera,
                           vacc_scenario = vacc_scenario,
                           year = eval(parse(text = year_chr)),
                           life_expectancy = life_exp,
                           parameter_sample = params[i,])

  cohort[[i]] <- res$cohort_size
  cases[[i]] <- res$cases
  deaths[[i]] <- res$deaths
  dalys[[i]] <- res$dalys
}
cases_sum <- Reduce('+', cases)
cases_sum_col <- colSums(cases_sum)
plot(2000:2100, cases_sum_col/nruns, type="l")
max(cases_sum_col/nruns)
summary(cases_sum_col/nruns)
```


## Simulation
Some methods require disease and other don't. How do I streamline it?
```{r}
# Check lists
# Are country names consistent across data sets?
# mean, min, mix functions na.rm = TRUE argument necessary?
# generate parameter sets to explore
# run vaccine impact
devtools::load_all()
set.seed(1234)
nruns <- 30
#VIMC report requires capital T in Typhoid
dis <- "Cholera"
year_chr <- "2000:2100"
vaccination_scenarios <- c("novacc", "campaign")
vaccination_scenarios_save <- c("novacc", "campaign")

tstamp <- format(Sys.time(), "%Y%m%dT%H%M")
# # target countries
# template <- 
#   fread("inst/extdata/central-burden-template.202110gavi-2.Cholera_IVI-Kim_standard.csv")
# countries <- clean_country_names(unique(template$country_name))
countries <- target_countries_cholera
# parameters to explore
# library(magrittr)
parameter_data <- parameters %>% filter(tolower(disease) == tolower(dis))

# Sampling based on Sobol low discrepancy sequence
params_list <- list()
for (k in 1:length(countries)) {
  cntry <- countries[[k]]
  params_list[[k]] <-
    set_parameter_cholera(disease = dis,
                          nruns = nruns,
                          country = cntry,
                          parameter_data = parameter_data,
                          incidence_rate_data = overall_incid_rate_cholera,
                          case_fatality_ratio_data = cfr_data)
}

saveRDS(params_list, paste0("outputs/params_", dis, "_", tstamp, ".rds"))
params_list <- readRDS(paste0("outputs/params_", dis, "_", tstamp, ".rds"))

params_list <- readRDS("outputs/params_Cholera_20211125T0734.rds")

stoch_output <- list()
stoch_parameters <- list()
central_output <- list()

# one vaccine coverage input
# coverage_202110gavi-2_cholera-no-vaccination.csv
# coverage_202110gavi-2_cholera-campaign-default.csv
# vacc_cov_input_campaign <- 
#   fread("inst/extdata/coverage_202110gavi-2_cholera-campaign-default.csv")
# vacc_cov_input_campaign$country <- clean_country_names(vacc_cov_input_campaign$country)

vacc_cov_input <- list()
vacc_cov_input[[1]] <- vacc_cov_input_cholera

## arrange vacc_scenarios according to the vaccination scenarios
vacc_cov_data <- vacc_cov_input[[1]]

# r = 2
# k = 1
# for (r in seq_along(vaccination_scenarios)) {
for (r in 1:1) {
  vacc_scenario <- vaccination_scenarios[r]
  vacc_scenario_save <- vaccination_scenarios_save[r]
  if (r > 1) {
    vacc_cov_data <- vacc_cov_input[[r - 1]]
  }
  for (k in seq_along(countries)) {
    cntry <- countries[k]
    cohort <- list()
    cases <- list()
    deaths <- list()
    dalys <- list() 

    life_exp <- 
      get_life_expectancy(life_expectancy_data = life_expectancy_data,
                          country = cntry,
                          year = eval(parse(text = year_chr)))
    
    message(paste0("vaccine scenario = ", 
                   vacc_scenario, ", country = ", cntry, 
                   ", ", k, "/", length(countries)))
    ## country-specific parameters
    params <- params_list[[k]] # country by index
    # i = 1
    for (i in 1:nruns) {
      # message(paste0("r = ", r, ", k = ", k, ", i = ", i))

      res <- vaccine_impact(
          disease = dis,
          country = cntry,
          population = population_data,
          wash_prop = wash_prop,
          wash_risk_ratio = wash_risk_ratio_chol,
          vacc_scenario = vacc_scenario,
          vaccine_coverage = vacc_cov_data,
          year = eval(parse(text = year_chr)),
          life_expectancy = life_exp,
          incidence_rate = params$incidence_rate[i],
          disability_weight = params$disability_weight[i],
          case_fatality = params$case_fatality_ratio[i],
          illness_duration = 5/365, 
          vaccine_immunity_duration = NULL,
          exponential_decay = FALSE,
          vaccine_efficacy = params$ve_yr0[i],
          vaccine_efficacy_by_year = c(
            params$ve_yr0[i],
            params$ve_yr1[i],
            params$ve_yr2[i],
            params$ve_yr3[i],
            params$ve_yr4[i],
            params$ve_yr5[i],
            params$ve_yr6[i]))
      
      cohort[[i]] <- res$cohort_size
      cases[[i]] <- res$cases
      deaths[[i]] <- res$deaths
      dalys[[i]] <- res$dalys 
    }
    
    stoch <- list()
    stoch$var <- c("cohort_size", "cases", "deaths", "dalys")
    stoch$val[[1]] <- cohort[[1]] # cohort size remains constant across nruns
    stoch$val[[2]] <- cases
    stoch$val[[3]] <- deaths
    stoch$val[[4]] <- dalys
    
    stoch_output[[k]] <- 
      vimc_stoch_report(disease = dis, country = cntry, sim_output = stoch)
    
    # stoch_parameters[[k]] <- stoch_param
    
    names(stoch$val) <- stoch$var
    central_output[[k]] <- vimc_central_report(disease = dis, country = cntry, 
                               sim_output = stoch$val)
  }
  saveRDS(stoch_output, paste0("outputs/stoch_output_list_", tstamp, "_", dis, "_", vacc_scenario_save, ".rds"))
   saveRDS(central_output, paste0("outputs/central_output_list_", tstamp, "_", dis, "_", vacc_scenario_save, ".rds"))
}
```


## Combine results in list into a single data.frame
Results across countries given vaccination scenario (i.e., novacc, campaign, or
routine) are in a list type. The do.call(rbind, x) is applied to make them a 
single data.frame   

### Manual single file conversion 
```{r}
# data in which results from each country is stored in a list
res_list <- 
  readRDS("outputs/central_output_list_20211222T212131_Cholera_campaign.rds")

# bind each element of the list by row
res <- do.call(rbind, res_list)
dis <- "Cholera" 
tstamp <- "20211222T212131"
# vacc <- c("novacc")
vacc <- c("campaign")

saveRDS(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".rds"))
fwrite(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".csv"))
```

### Auto-conversion across vaccination scenarios
```{r}
disease <- c("Cholera")
vacc_scenario <- c("novacc", "campaign")
tstamp <- "20211221T00"
library(data.table)
# vacc <- vacc_scenario[2]
for(dis in disease) {
  for (vacc in vacc_scenario) {
    # central <- list.files("outputs", pattern = paste0("central.*list.*", tstamp, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)
    # ## bind by row from a list of data with the same format
    # res <- do.call(rbind, readRDS(central))
    # saveRDS(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".rds"))
    # fwrite(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".csv"))
    
    stoch_output <- list.files("outputs", pattern = paste0("stoch.*output.*list.*", tstamp, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)
    res_stoch <- do.call(rbind, readRDS(stoch_output))
    saveRDS(res_stoch, paste0("outputs/stoch_", dis, "_", vacc, "_", tstamp, ".rds"))
    fwrite(res_stoch, paste0("outputs/stoch_", dis, "_", vacc, "_", tstamp, ".csv"))
    }
}
```


### Effective vaccine coverage
```{r}
dis <- "Cholera"

cntry <- target_countries_cholera[1]
for( cntry in target_countries_cholera) {
  pop <- setup_cohorts(country = cntry,
                       year = 2000:2100,
                       population_data = population_data)
  vacc_cov <-
    create_vaccine_coverage(country = cntry,
                            population = pop,
                            vaccine_coverage_data = vacc_cov_input_cholera)
  
  params_list <- readRDS("outputs/params_Cholera_20211220T130658.rds")
  parameter_sample <- params_list[[1]] # country by index
  
  vacc_eff_by_year <- unlist(parameter_sample[paste0("vacc_eff_year_", 0:6)])
      
  vacc_protected <-
    calculate_vaccine_protected(disease = dis,
                                country = cntry,
                                population = pop,
                                vaccine_coverage = vacc_cov,
                                vaccine_efficacy = NULL,
                                vaccine_immunity_duration = NULL,
                                exponential_decay = FALSE,
                                vaccine_efficacy_by_year = vacc_eff_by_year)
  vacc_recipients <-
    calculate_vaccine_protected(disease = dis,
                              country = cntry,
                              population = pop,
                              vaccine_coverage = vacc_cov,
                              vaccine_efficacy = 1.0,
                              vaccine_immunity_duration = 1000,
                              exponential_decay = TRUE,
                              vaccine_efficacy_by_year = NULL)
  
  pop_unprotected <- pop
  # factor protection for vaccine recipients
  pop_unprotected <- pop_unprotected - vacc_protected
  pop_unvacc <- pop - vacc_recipients
      # effective vaccine coverage used to switch on/off indirect effect
  eff_vacc_cov_year <-
    calc_vacc_cov_year(pop = pop, vacc_recipients = vacc_protected)
  
  vacc_cov_year <-
    calc_vacc_cov_year(pop = pop, vacc_recipients = vacc_recipients)
  
  IVE <- calc_indirect_vacc_protected_cholera(population = pop_unvacc, eff_vacc_cov = eff_vacc_cov_year) 
  propIVE <- colSums(IVE) / colSums(pop)
  plot(propIVE, type="l", main=cntry, ylim=c(0, 1))
  lines(eff_vacc_cov_year, col=2)
  lines(vacc_cov_year, col=3)
  invisible(readline(prompt = "Press RET to continue"))
}
```

### Compare w/ and w/o indirect effects (edit)
```{r}
# sim1 <- readRDS("outputs/central_Cholera_campaign_20211220T175515.rds")
sim1 <- readRDS("outputs/central_Cholera_campaign_20211221T004050.rds")
sim2 <- readRDS("outputs/central_Cholera_campaign_20211222T212131.rds")

# cntry = target_countries_tf[1]
dis = "Cholera"

library(ggplot2)
library(gridExtra)
for (cntry in target_countries_cholera){
  plt1 <- year_val_plot(disease = dis, country = cntry, 
                        sim_central_value = sim1)
  plt2 <- year_val_plot(disease = dis, country = cntry, 
                        sim_central_value = sim2)
  grid.arrange(plt1, plt2, nrow=2)
  invisible(readline(prompt = "Press RET to continue"))
}
```

## Check values
```{r}
cntry <- "Afghanistan"
for (cntry in target_countries_cholera){
  message(cntry)
  sim1 %>%
      filter(country_name == cntry) %>%
      group_by(year) %>%
      summarize(cohort_size = sum(cohort_size),
                cases = sum(cases),
                deaths = sum(deaths),
                dalys = sum(dalys))
    invisible(readline(prompt = "Press RET to continue")) }
```

## Diagnostic plots
```{r}
library(ggplot2)
# tstamp <- "20211220T134753"
# tstamp <- "20211125"
tstamp <- "20211220T185451"
# central_Cholera_novacc_20211220T185451.rds
# central_Cholera_novacc_20211125.rds
for (cntry in target_countries_cholera) {
  for(tstamp in c("20211125", "20211220T185451")) {
    dis <- "Cholera"
    vacc <- "novacc"
    # cntry <- "Iran, Islamic Republic of"
    res <- readRDS(paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".rds"))
    plt <- year_val_plot(disease = dis, country = cntry, sim_central_value = res)
    # year_val_plot(disease = dis, country = cntry, sim_central_value = res, age = 10)
    print(plt)
    invisible(readline(prompt = "Press RET to continue"))
  } 
}
```

## Diagnostic plots II
Visualize cohort size, cases, deaths, and dalys across vaccination scenarios for each of the countries over the period of 2000 - 2100
```{r}
library(ggplot2)
library(grid)
library(gridExtra)
dis <- "Cholera"
tstamp <- "20211125"
for (cntry in countries){
  res1 <- readRDS(paste0("outputs/central_",  dis, "_", "novacc_", tstamp, ".rds"))
  res2 <- readRDS(paste0("outputs/central_",  dis, "_", "campaign_", tstamp, ".rds"))
  
  p1 <- year_val_plot(disease = dis, country = cntry, sim_central_value = res1, vacc_scenario = "novacc")
  p2 <- year_val_plot(disease = dis, country = cntry, sim_central_value = res2, vacc_scenario = "campaign")

  plt <- grid.arrange(p1, p2, ncol = 1)
  print(plt)
  invisible(readline(prompt = "Press RET to continue"))
} 
```

## Diagnostic III
1. Check if there are population categories (defined by chronological year and age) have negative numbers
2. Check if cohort sizes are same across vaccination scenarios 
3. Check if cases, deaths, and dalys are larger for non-vaccination scenarios compared to vaccination scenarios
3. Check if cases, deaths, and dalys are larger for routine (or campaign) scenario compared to the 
```{r}
dis <- "Cholera"
tstamp <- "20211125"
for (cntry in countries){
  res1 <- readRDS(paste0("outputs/central_",  dis, "_", "novacc_", tstamp, ".rds"))
  res2 <- readRDS(paste0("outputs/central_",  dis, "_", "campaign_", tstamp, ".rds"))
  
  if (sum(res1$cohort_size != res2$cohort_size) != 0) {
    cat("Cohort size of novacc and routine scenrios are different for", cntry, "\n")
  }
  if (sum(res1$cases < res2$cases) != 0) {
    cat("Cases are bigger for routine scenario than novacc scenrios for", cntry, "\n")    
  }
   if (sum(res1$deaths < res2$deaths) != 0) {
    cat("Deaths are bigger for routine scenario than novacc scenrios for", cntry, "\n")    
  }
  if (sum(res1$dalys < res2$dalys) != 0) {
    cat("DALYs are bigger for routine scenario than novacc scenrios for", cntry, "\n")
  }
}
```

# Diagnostic IV: Negative population size
Negative values can happen when there is population size for a certain age group
is zero and in the previous year there are vaccinated people in one-year younger
age group some value and model predicts some people move to the next age group,
which is determined by the probability of dying and aging, vaccine  

```{r}
# vacc_scenario <- c("novacc", "campaign")
vacc_scenario <- c("campaign")
tstamp <- "20211221T004050"

for (vacc in vacc_scenario) {
  fl <- paste0("outputs/central_Cholera_", vacc, "_", tstamp, ".csv")
  message(paste0("file = ", fl))
  d <- fread(fl)
  if(sum(d$cohort_size < 0) > 0){
    message("Negative cohort size!")
  }
  if(sum(d$cases < 0) > 0){
    message("Negative cases!")
  }
  if(sum(d$deaths < 0) > 0){
    message("Negative deaths!")
  }
  if(sum(d$dalys < 0) > 0){
    message("Negative dalys!")
  }
}
```

## Prepare csv files for uploading the results onto Montagu
```{r}
library(data.table)
dis <- "Cholera"
tstamp <- "20211125"

for (vacc in c("novacc", "campaign")) {
  fl <- list.files("outputs/", pattern = paste0("central.*", dis, ".*", vacc, ".*", tstamp,  ".*rds$"), full.names = T)
  d <- readRDS(fl)
  fwrite(d, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".csv"))
}

# fls <- list.files("outputs", pattern = paste0("^param_combined.*", tstamp, ".*", dis, ".*", ".*rds$"), full.names = T)
# x <- lapply(fls, function(x) readRDS(x))
# d <- do.call(rbind, x)
# fwrite(d, paste0("outputs/param_", dis, "_combined", tstamp, ".csv"))
```


## Combine parameters
```{r}
params_list <- readRDS("outputs/params_Cholera_20211220T130658.rds")
# modify the list to 
# lst <- list()
# for (i in 1:length(target_countries)) {
#   params <- params_list[[i]]
#     ir <- params[[1]]
#   t_ir <- data.table::transpose(ir)
#   colnames(t_ir) <- paste0("incid_rate_ag_", 0:100)
#   cfr <- params[[2]]
#   t_cfr <- data.table::transpose(cfr)
#   colnames(t_cfr) <- paste0("case_fatality_ratio_ag_", 0:100)
#   lst[[i]] <- cbind(params[[3]], t_ir, t_cfr)
# }
# res <- do.call(rbind, lst)

res <- do.call(rbind, params_list)
# run_id has to be the first column and therefore, files are edited, 
# set_params_cholera was also adjusted and therefore will not be necceary for 
# future runs
res$disease <- "Cholera"
dis <- "Cholera"
res <- res[, c(2, 18, 1, 3:17)]
names(res)
tstamp <- "20211221T00" # use the same time stamp as the simulation results
saveRDS(res, paste0("outputs/params_", dis, "_", tstamp, ".rds"))
fwrite(res, paste0("outputs/params_", dis, "_", tstamp, ".csv"))
```

