---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


### Mean age of infection 
Mean age of infection, $A$, can be used to approximate $R_0$ for a given 
life expectancy, $L$. 
$$R_0 \approx L/A$$
calculating $A$ can be done using force of infection, $\lambda$

```{r}
# force of infection, lambda, can be interpreted as the probability of infection per year
lambda <- 0.012 # per year per person
# assuming exponential distribution and infinite time makes the average waiting time is 1 / lambda
A <- 1 / lambda
A
```

### What happens if we have a limited life span (i.e., life expectancy)?
```{r}
# force of infection, lambda, can be interpreted as the probability of infection per year
lambda <- 0.012 # per year per person
L <- 80 # life expectancy
# mean age of infection is the same as the proportion of people remain susception 
A <- 1/(lambda + 1/L)
A
```

### Age distribution by year 
```{r}
pop_by_age = c(120, 99, 78, 60)
foi_by_age = c(0.01, 0.1, 0.05, 0.01)
age = 0:3
age_inf = age + (foi_by_age + 1)
A = sum(age_inf * pop_by_age / sum(pop_by_age))
A
```

### ODE model - SIRS 
```{r}
ode_open_sir <- function(t, y, params) {
  S <- y[1]; I <- y[2]; R <- y[3]; CI <- y[4]; FOI <- y[5]  ## now extract the parameters
  beta <- params["beta"]
  gamma <- params["gamma"]
  mu <- params["mu"]

  N <- S + I + R # total population
  FOI <- beta * I / N # force of infection
  
  dS <- - FOI*S + mu*(N-S)
  dI <- FOI*S - gamma*I - mu*I
  dR <- gamma*I - mu*R
  dCI <- FOI*S
  dFOI <- FOI
  
  return(list(c(dS, dI, dR, dCI, dFOI)))
}
```

## Check $A, L, R_0$ using the ODE model

```{r}
library(deSolve)
library(tidyverse)
L = 20 # life expectancy, which is intentionally set to be short to make the dynamics converge quickly
R0 = 1.5
gamma = 0.2*365
beta = R0 * (gamma + 1/L)

params <- c(beta = beta, gamma = gamma, mu = 1/L) # per year
y0 <- c(S = 1e5-10, I = 10, R = 0, CI = 0, FOI = 0)
tend <- 500
times <- seq(0, tend, by = 1) # daily output for 150 days

ode(y = y0, times = times, func = ode_open_sir, parms = params) %>% 
  as.data.frame() -> out 

out %>% 
  ggplot(aes(time, FOI)) +
  geom_line()

lambda = tail(diff(out$FOI), 1)
A = 1 / (lambda + 1/L)
A
L/A
```

## Two-age group model
```{r}
ode_open_sir_2ag <- function(t, y, params) {
  S_1 <- y[1]; I_1 <- y[2]; R_1 <- y[3]; CI_1 <- y[4]; FOI_1 <- y[5]  
  S_2 <- y[6]; I_2 <- y[7]; R_2 <- y[8]; CI_2 <- y[9]; FOI_2 <- y[10]
  
  beta <- params["beta"]
  gamma <- params["gamma"]
  mu <- params["mu"]
  ag <- params["ag"]

  N <- S_1 + I_1 + R_1 + S_2 + I_2 + R_2 # total population
  FOI <- beta * (I_1 + I_2) / N # force of infection
  FOI_2 <- FOI_1 <- FOI
  
  dS_1 <- - FOI*S_1 + mu*(N-S_1) - ag*S_1
  dI_1 <- FOI*S_1 - gamma*I_1 - mu*I_1 - ag*I_1
  dR_1 <- gamma*I_1 - mu*R_1 - ag*R_1
  dCI_1 <- FOI*S_1
  dFOI_1 <- FOI_1
  
  dS_2 <- - FOI*S_2 - mu*S_2 + ag*S_1 
  dI_2 <- FOI*S_2 - gamma*I_2 - mu*I_2 + ag*I_1
  dR_2 <- gamma*I_2 - mu*R_2 + ag*R_1
  dCI_2 <- FOI*S_2
  dFOI_2 <- FOI_2
  
  return(list(c(dS_1, dI_1, dR_1, dCI_1, dFOI_1, dS_2, dI_2, dR_2, dCI_2, dFOI_2)))
}
```

## Check $A, L, R_0$ using the ode model

```{r}
library(deSolve)
library(tidyverse)
## R0 = 1.5 and L = 10
L = 20 # life expectancy, which is intentionally short to make the dynamics converge quickly
mu = 1 / L
R0 = 1.2
gamma = 0.2*365
beta = R0 * (gamma + mu)
ag = 1/2

params <- c(beta = beta, gamma = gamma, mu = mu, ag = ag) # per year
y0 <- c(S_1 = (1e5-10)/2, I_1 = 10/2, R_1 = 0, CI_1 = 0, FOI_1 = 0,
        S_2 = (1e5-10)/2, I_2 = 10/2, R_2 = 0, CI_2 = 0, FOI_2 = 0)
tend <- 500
times <- seq(0, tend, by = 1) # daily output for 150 days

ode(y = y0, times = times, func = ode_open_sir_2ag, parms = params) %>% 
  as.data.frame() -> out 

out %>% 
  ggplot(aes(time, FOI_1)) +
  geom_line()

out %>% 
  ggplot(aes(time, S_1)) +
  geom_line()

lambda = tail(diff(out$FOI_1), 1)

S_1 = tail(out$S_1, 1)
S_2 = tail(out$S_2, 1)

out_rates_ag1 <- ag + mu + lambda

A_ag = S_1/(S_1 + S_2)*(1/out_rates_ag1) + 
  S_2/(S_1 + S_2)*(1/out_rates_ag1 + 1/(lambda + mu))

A = 1 / (lambda + 1/L)
A
A_ag
L/A
L/A_ag
```

I think a better strategy would be to calculate averaged $lambda$ and then calculate the $L/A$ where $L$ should come from the final age group

$$\omega = \frac{F(p)}{F(0)} = \frac{R_0(1-Pe)-1}{R_0 - 1}$$

## Test with the simulation data set
```{r}
devtools::load_all()

library(tidytable)
library(tidyverse)

template <- 
  fread("inst/extdata/central-burden-template.202108test-1.Typhoid_IVI-Kim_standard.csv")
target_countries <- clean_country_names(unique(template$country_name))
cntry <- target_countries[1]
# parameters to explore
parameter_data <- fread("inst/extdata/parameters.csv")
vacc_cov_data <- 
  fread('inst/extdata/coverage_202108test-1_typhoid-routine-default.csv')
population_data <- as.data.frame(fread("outputs/population_wide.csv"))

cntry <- unique(vacc_cov_data$country)[1]
incid_data_typhoid %>%
  filter.(measure == "Incidence", year == 2015, location %in% cntry) %>%
  as.data.frame() -> ir_data

ir <- spread_age_IHME(ir_data)

year <- "2000:2100"
pop <- setup_cohorts(country = cntry,
                     year = eval(parse(text = year)),
                     population_data = population_data)

pop_wt <- apply(pop, 2, function(x) x / colSums(pop))
foi <- rep(NA, ncol(pop_wt))
for(i in 1:ncol(pop_wt)) {
  foi[i] <- sum(ir/1e5 * pop_wt[, i])
}

life_expectancy <- 66.6 # Ethiopia
A <- 1 / (foi + 1/life_expectancy)
R0 <- life_expectancy / A
summary(R0)

calc_mean_R0 <- function(L) {
  mean_R0 <- rep(NA, length(L))
  for (i in 1:length(L)) {
    A <- 1 / (foi + 1/L[i])
    mean_R0[i] <- mean(L[i] / A)
  }
  return(mean_R0)
}

L <- 50:100
plot(L, calc_mean_R0(L))

foi[1]*L[1]/ (1 - exp(-foi[1]*L[1]))
(foi[1] + 1/L[1])*L[1]

A <- 1/(foi[6] + 1)

a <- 1/(foi + 1)
a <- a + 0:100
A <- sum(a * pop_wt[,20])
```

## vaccine coverage
```{r}
devtools::load_all()

library(tidytable)
library(tidyverse)

template <- 
  fread("inst/extdata/central-burden-template.202108test-1.Typhoid_IVI-Kim_standard.csv")
target_countries <- clean_country_names(unique(template$country_name))
cntry <- target_countries[1]
# parameters to explore
parameter_data <- fread("inst/extdata/parameters.csv")
vacc_cov_data <- 
  fread('inst/extdata/coverage_202108test-1_typhoid-routine-default.csv')
population_data <- as.data.frame(fread("outputs/population_wide.csv"))

cntry <- unique(vacc_cov_data$country)[1]
incid_data_typhoid %>% 
  filter.(measure == "Incidence", year == 2015, location %in% cntry) -> ir_data

ir <- spread_age_IHME(as.data.frame(ir_data))

year <- "2000:2100"
pop <- setup_cohorts(country = cntry,
                     year = eval(parse(text = year)),
                     population_data = population_data)

vacc_cov <- create_vaccine_coverage(country = cntry,
                                    population = pop,
                                    vaccine_coverage_data =        
                                      vaccine_coverage_data)

pop_wt <- apply(pop, 2, function(x) x / colSums(pop))

overall_cov <- rep(NA,  ncol(pop_wt))
for (i in 1:ncol(pop_wt)) {
  overall_cov[i] <- sum(vacc_cov[, i] * pop_wt[, i])
}
summary(overall_cov)
```


```{r}

library(ggplot2)
library(manipulate)

R0 = 1.1
calc_omega <- function(R0, Pe){
  omega = (R0*(1-Pe)-1)/(R0 - 1)
  omega[omega < 0] = 0
  return (omega)
}

manipulate(
  { 
  Pe = seq(0.01, 0.3, length.out = 50)
  omega = calc_omega(R0, Pe)
  data.frame(Pe = Pe,
             omega = omega) %>% 
  ggplot(aes(Pe, omega)) +
    geom_line() +
    labs(x = "Pe", y = "Adjustment factor, omega", color = "")
  },
  R0 = slider(1.1, 1.5, initial = 1.1, step = 0.05)
)

```

```{r}
  output$cohort_size <- pop # cohort size is assumed to be the total population


  if (tolower(vacc_scenario) %in% c("routine", "campaign")) {
    vacc_protected <-
      calculate_vaccine_protected(
        disease = dis,
        country = cntry,
        population = pop,
        vaccine_coverage = vacc_cov,
        vaccine_efficacy = parameter_sample$vacc_efficacy,
        vaccine_immunity_duration = parameter_sample$duration_vacc_protection,
        exponential_decay = TRUE,
        vaccine_efficacy_by_year = NULL)

    pop <- pop - vacc_protected
}
```

