---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Read and save data
```{r}
library(data.table)
library(devtools)
library(usethis)
library(tidyverse)

# incidence_rate <- fread("data/incidence_rate.csv")
# incidence_rate[, country := clean_country_names(country)] 
# # incidence_rate["country"] <- lapply(incidence_rate["country"], function(x) clean_country_names(x))
# # add cholera data
# cholera_ir <- fread("data/cholera_incidence.csv")
# cholera_ir$Country <- clean_country_names(cholera_ir$Country)
# names(cholera_ir) <- c("disease", "country", "incidence_rate_100Kpyo") # to match the names of incidence_rate
# 
# ir_ty <- incidence_rate[disease == "typhoid", c("disease", "country", "incidence_rate_100Kpyo")]
# incidence_rate <- rbind(ir_ty, cholera_ir)
# incidence_rate[disease == "typhoid", disease := "Typhoid"]
# incidence_rate$country <- clean_country_names(incidence_rate$country)
# 
# use_data(incidence_rate, overwrite = T)
  
# incidence_rate[, country := clean_country_names(country)] 

# for (i in 1:nrow(gavi201910_int_pop_both)){
#   incidence_rate$country[i] <- clean_country_names(incidence_rate$country[i])
# }
## to make the country name consitent: DR Congo -> "Congo, the Democratic Republic of the"
# incidence_rate %>% filter(country != "DR Congo") -> ir1
# incidence_rate %>% filter(country == "DR Congo") %>% mutate(country = "Congo, Democratic Republic of the") -> ir2
# incidence_rate <- bind_rows(ir1, ir2)

# gavi201910_int_pop_both <- fread("data/201910gavi-5_dds-201910_2_int_pop_both.csv")
# gavi201910_int_pop_both[, country := clean_country_names(country)] 
# use_data(gavi201910_int_pop_both, overwrite = TRUE)
# 
# life_expectancy <- fread(life_expectancy_data)
# life_expectancy[, country := clean_country_names(country)] 
# use_data(life_expectancy, overwrite = TRUE)
# 
# prop_basic_san <- setDT(prop_basic_san)
# prop_basic_san[, country := clean_country_names(country)] 
# use_data(prop_basic_san, overwrite = TRUE)
# 
# parameters <- fread("data/parameters.csv")
# usethis::use_data(parameters, overwrite = T)
# 
p_dying_age <- fread("data/201910gavi-5_dds-201910_p_dying_both.csv")
p_dying_age[, country := clean_country_names(country)] 
use_data(p_dying_age, overwrite = T)
# 
# country_name_table <- fread("data/country_names.csv")
# names(country_name_table) <- c("country", "iso3")
# country_name_table <- country_name_table[1:112,] # remove the bottom row, Total
# country_name_table[, country := clean_country_names(country)] 
# use_data(country_name_table, overwrite = T)
```

## Population at risk of cholera or typhoid fever
We assume that people who have at least basic sanitation are safe from these diseases (or at least, have reduced risk).
We used the WaSH data from JMP (2000-2017) to fit a saturating exponential function to roughly estimate the trend of WaSH.

```{r}
# d <- fread("data/JMP2019.csv")
# d <- d[-(1:3),]
# d <- d %>% filter(V3 == 'National') %>% select(V1, V2, V3, V17) # country, year, rural/urban/national, percent population with access to basic sanitation services
# names(d) <- c("country", "year", "rural/urban/national", "prop_san")
# d$year <- as.double(d$year)
# d$prop_san <- as.double(d$prop_san)/100 # make percents to fractions 
# d
# san <- d %>% filter(country == "Congo, Democratic Republic of the")
# # san <- d %>% filter(country == 'Ethiopia')
# # san <- d %>% filter(country == 'Democratic Republic of the Congo')
# # san <- d %>% filter(country == 'India')
# # san <- d %>% filter(country == 'Nigeria')
# 
# f <- function (theta, x) {1 - exp(-theta[2] * (x - theta[1]))} # saturating exponential function is fitted
# 
# yr <- san$year
# prop <- san$prop_san
# 
# ssq <- function(theta){
#   tau <- theta[1]
#   k <- theta[2]/(1-theta[2])
#   sum((prop - (1 - exp(-k * (yr - tau))))^2)
# }
# 
# library(optimx)
# start <- c(1990, 1e-3)
# lower <- c(1e-6, 1e-6)
# upper <- c(2000-1e-3, 10)
# 
# fit <- optimx(par = start,
#               fn = ssq,
#               method = 'nlminb',
#               lower = lower, 
#               upper = upper,
#               itnmax = 100,
#               control = list(trace=5))
# 
# x <- 2000:2100 
# plot(x, 1 - exp(-fit$p2 * (x - fit$p1)), type='l')
# points(yr, prop, col=2)

# prop_san <- data.frame(country = "Pakistan", yr = 2000:2100)
# prop_san$pred <- 1 - exp(-fit$p2 * (prop_san$yr - fit$p1))
# 
# prop_san2 <- data.frame(country = "Ethiopia", yr = 2000:2100)
# prop_san2$pred <- 1 - exp(-fit$p2 * (prop_san2$yr - fit$p1))

# prop_san3 <- data.frame(country = "Democratic Republic of the Congo", yr = 2000:2100)
# prop_san3$pred <- 1 - exp(-fit$p2 * (prop_san3$yr - fit$p1))

# prop_san4 <- data.frame(country = "India", yr = 2000:2100)
# prop_san4$pred <- 1 - exp(-fit$p2 * (prop_san4$yr - fit$p1))

# prop_san5 <- data.frame(country = "Nigeria", yr = 2000:2100)
# prop_san5$pred <- 1 - exp(-fit$p2 * (prop_san5$yr - fit$p1))

# prop_basic_san <- rbind(prop_san, prop_san2, prop_san3, prop_san4, prop_san5)
# ## Change "Democratic Republic of the Congo" to "Congo, the Democratic Republic of the"
# prop_basic_san %>% filter(country != "Democratic Republic of the Congo") -> p1
# prop_basic_san %>% filter(country == "Democratic Republic of the Congo") %>% mutate(country = "Congo, the Democratic Republic of the") -> p2
# prop_basic_san <- bind_rows(p1, p2)
# names(prop_basic_san) <- c("country", "year", "prop")
# use_data(prop_basic_san, overwrite = TRUE)
```

## Estimate the proportion of population with access to basic sanitation services
WaSH data were downloaded from [JMP](https://washdata.org/data/downloads#WLD) and Household World file.
The data in Excel worksheet format has 4 tabs: introduction and 3 data sheets (water, sanitation, and hygiene).
We assume that people with at least basic sanitation are protected from cholera, which is a simplifying assumption, which will be addressed (eg, we can alternatively assume that they have lower risk of infection). 
The data file was opened in MS Excel and each data tab was saved as csv (JMP2019_WLD_water.csv, JMP2019_WLD_sanitation.csv, JMP2019_WLD_hygiene.csv) 

```{r}
# d <- fread("data/JMP_2019_WLD_sanitation.csv")
# d <- d[-c(1,2,4179,4180), c(1,3,6)] # after viewing the file, rows and columns were decided.
# names(d) <- c("country", "year", "prop_san") # proportion of population with at least basic sanitation 
# d[, year := as.double(year)]
# d[, prop_san := as.double(prop_san)/100]# make percents to fractions 
# d <- d[!is.na(d$prop_san),]
# d[, country := clean_country_names(country)]
# 
# ## Select a country of interest
# san <- d %>% filter(country == target_countries_cholera[5])
# 
# f <- function (theta, x) {1 - exp(-theta[2] * (x - theta[1]))} # saturating exponential function is fitted
# 
# yr <- san$year
# prop <- san$prop_san
# 
# ssq <- function(theta){
#   tau <- theta[1]
#   k <- theta[2]/(1-theta[2])
#   sum((prop - (1 - exp(-k * (yr - tau))))^2)
# }
# 
# library(optimx)
# start <- c(1990, 1e-3)
# lower <- c(1e-6, 1e-6)
# upper <- c(2000-1e-3, 10)
# 
# fit <- optimx(par = start,
#               fn = ssq,
#               method = 'nlminb',
#               lower = lower, 
#               upper = upper,
#               itnmax = 100,
#               control = list(trace=5))
# 
# ## Quick diagnostic plot
# x <- 2000:2100 
# plot(x, 1 - exp(-fit$p2 * (x - fit$p1)), type='l')
# points(yr, prop, col=2)
# 
# prop_san <- data.frame(country = target_countries_cholera[2], yr = 2000:2100)
# prop_san$pred <- 1 - exp(-fit$p2 * (prop_san$yr - fit$p1))
# # 
# prop_san2 <- data.frame(country = target_countries_cholera[3], yr = 2000:2100)
# prop_san2$pred <- 1 - exp(-fit$p2 * (prop_san2$yr - fit$p1))
# 
# prop_san3 <- data.frame(country = target_countries_cholera[4], yr = 2000:2100)
# prop_san3$pred <- 1 - exp(-fit$p2 * (prop_san3$yr - fit$p1))
# 
# prop_san4 <- data.frame(country =  target_countries_cholera[5], yr = 2000:2100)
# prop_san4$pred <- 1 - exp(-fit$p2 * (prop_san4$yr - fit$p1))
# 
# prop_basic_san2 <- rbind(prop_san, prop_san2, prop_san3, prop_san4)
# names(prop_basic_san2) <- c("country", "year", "prop")
# prop_basic_san <- rbind(prop_basic_san, prop_basic_san2)
# use_data(prop_basic_san, overwrite = TRUE)
```

## Vaccine coverage data
Data files in csv format were downloaded from the VIMC modeling through the following steps:
1.	Log on at https://montagu.vaccineimpact.org/
2.	Select ‘Modellers Contribution Portal’
3.	Select ‘201910gavi (version 5)’

```{r}
# vacc_cov_input_campaign_cholera <- fread("data/coverage_201910gavi-5_cholera-campaign-default-test.csv") 
# vacc_cov_input_campaign_cholera$country <- clean_country_names(vacc_cov_input_campaign_cholera$country)
# vacc_cov_input_routine_typhoid <- fread("data/coverage_201910gavi-5_typhoid-routine-default-test.csv")
# vacc_cov_input_campaign_typhoid <- fread("data/coverage_201910gavi-5_typhoid-campaign-default-test.csv")
# vacc_cov_input_routine_typhoid$country <- clean_country_names(vacc_cov_input_routine_typhoid$country)
# vacc_cov_input_campaign_typhoid$country <- clean_country_names(vacc_cov_input_campaign_typhoid$country)
# 
# usethis::use_data(vacc_cov_input_routine_typhoid, overwrite = T)
# usethis::use_data(vacc_cov_input_campaign_typhoid, overwrite = T)
# usethis::use_data(vacc_cov_input_campaign_cholera, overwrite = T)
```

## Check functions
```{r}
devtools::load_all(".")

cntry <- "India"
dis <- "Typhoid"
pop <- setup_cohorts(country = cntry)
cases <- calculate_cases(population = pop, country = cntry, disease = dis)
yld <- calculate_YLD(cases = cases, disease = dis)
yll <- calculate_YLL(cases = cases, country = cntry, disease = dis)
daly <- calculate_DALY(cases = cases, country = cntry, disease = dis)
daly1 <- yll + yld
daly[1:5, 1:5]
daly1[1:5, 1:5]
# ## document the function
# devtools::document()
```

## Check Vaccine scenarios
```{r}
target_countries <- c('Pakistan', 'Ethiopia', 'Congo, Democratic Republic of the', 'India', 'Nigeria')
target_countries_cholera <- c("Congo, Democratic Republic of the", "Kenya", "Somalia", "South Sudan", "Ethiopia")
cntry <- "Congo, Democratic Republic of the"
dis <- "Cholera"
dis <- "Typhoid"
# for (cntry in target_countries) {
  cntry <- target_countries[4]
  vc <- create_vaccine_coverage(country = cntry, disease = dis, routine = F)
  pop <- setup_cohorts(country = cntry)
  pop_vacc <- calculate_vaccine_protected(country = cntry, population = pop, disease = dis, vacc_coverage = vc)


vc_cmp_chol <- lapply(target_countries_cholera, function(x) create_vaccine_coverage(country = x, disease ="Cholera", routine = F))
vc_cmp_ty <- lapply(target_countries, function(x) create_vaccine_coverage(country = x, disease ="typhoid", routine = F))


```

## Simulation 
```{r}
# fls <- list.files("data", pattern = paste0("*.typ", ".*rds$"), full.names = T)
# unlink(fls)
devtools::load_all(".")
cntries_typhoid <- c('Pakistan', 'Ethiopia', 'Congo, Democratic Republic of the', 'India', 'Nigeria')
cntries_cholera <- c("Congo, Democratic Republic of the", "Kenya", "Somalia", "South Sudan", "Ethiopia")
dis <- c("Cholera", "Typhoid")
dis <- "Typhoid"
vacc_scenario = c("novacc", "campaign", "routine")
# vacc_scenario = c("novacc", "campaign")
set.seed(0)
for (d in dis) {
  if (d == "Cholera") {
    cntries <- cntries_cholera
  } else {
    cntries <- cntries_typhoid
  }
  for (vs in vacc_scenario) {
    if (!(d == "Cholera" & vs == "routine")) {
      stoch_run_scenario(disease = d, country = cntries, vacc_scenario = vs)
    }
  }
}
```

## Combine results into a single file
```{r}
# v_scenario <- "routine"
disease <- c("Cholera", "Typhoid")
disease <- c("Typhoid")
vacc_scenario <- c("novacc", "campaign", "routine")
for(dis in disease) {
  for (vacc in vacc_scenario) {
    if (!(dis == "Cholera" & vs == "routine")) {
      fls <- list.files("outputs", pattern = paste0("^rep.*", dis, ".*", vacc, ".*rds$"), full.names = T)
      fls
      x <- lapply(fls, function(x) readRDS(x))
      res <- do.call(rbind, x)
      saveRDS(res, paste0("outputs/central_", dis, "_", vacc, ".rds"))
    }
  }
}

```

## Diagnostic plots
```{r}
library(ggplot2)
dis <- "Cholera"
vacc <- "campaign"
cntry <- "Somalia"
res <- readRDS(paste0("outputs/central_",  dis, "_", vacc, ".rds"))
year_val_plot(disease = dis, country = cntry, sim_central_value = res)
year_val_plot(disease = dis, country = cntry, sim_central_value = res, age = 10)
```

## Report for the stochastic outputs
```{r}
dis <- c("Cholera", "Typhoid")
dis <- c("Typhoid")
vacc_scenario = c("novacc", "campaign", "routine")

for (d in dis) {
  for (vs in vacc_scenario) {
    if (!(d == "Cholera" & vs == "routine")) {
      if (d == "Cholera") {
        cntry <- cntries_cholera
      } else {
        cntry <- cntries_typhoid
      }
      list1 <- list()
      for(i in seq_along(cntry)){
        output1 <- readRDS(paste0("outputs/stoch_output_", d, "_",  cntry[i], "_", vs, ".rds"))
        list1[[i]] <- vimc_stoch_report(disease = d, country = cntry[i], sim_output = output1)
      }
      res <- do.call("rbind", list1)
      saveRDS(res, paste0("outputs/stoch_", d, "_", vs, ".rds"))
    }
  }
}
```


## Uploading the results
```{r}
library(data.table)
# dis <- "typhoid"
# v_scenario <- "routine"
# fls <- list.files("./data", pattern=paste0("central.*", dis, ".*", v_scenario, ".*rds$"), full.names = T)
# d <- readRDS(fls)
# fwrite(d, paste0("outputs/", sim, "_", dis, "_", v_scenario, ".csv"))


dis <- c("Cholera", "Typhoid")
dis <- c("Typhoid")
vacc <- c("novacc", "campaign", "routine")
sims <- c("stoch", "central")

for (d in dis) {
  for (v in vacc) {
    for (sim in sims){
      if (!(tolower(d) == "cholera" & tolower(v) == "routine" )){
        fls <- list.files("outputs", pattern = paste0(sim, "_", d, "_", v, ".rds$"), full.names = T)
        fls
        res <- readRDS(fls)
        res$disease <- d
        fwrite(res, paste0("outputs/", sim, "_", d, "_", v, ".csv"))
      }
    }
  }
}

##
# fwrite(parameters, "outputs/parameters.csv")
# fwrite(incidence_rate, "outputs/incidence_rate.csv")

```

```{r}
sim <- "stoch"
# d <- "Cholera"
v <- "novacc"
fls <- list.files("outputs/", pattern = paste0(sim, "(.*)param(.*).rds$"), full.names = T)
for (fl in fls) {
  res <- readRDS(fl)
  fwrite(res, gsub("rds$", "csv", fl))
}
fls
c1 <- readRDS(fls[2])
c1$vacc_type <- "No vaccination"
c2 <- readRDS(fls[1])
c2$vacc_type <- "Campaign"
c <- rbind(c1,c2)
c$`disability weight` <- 0.202
c$`duration of illness` <- "5 days"
c$`case fatality ratio` <- 0.030
c$`duration of vaccine-induced immunity` <- "3 years"
c$`vaccine efficacy` <- 0.640
c$`vaccine efficacy (< 5 yo)` <- 0.300

fwrite(c, "outputs/stoch_param_Cholera.csv")

t1 <- readRDS(fls[4])
t1$vacc_type <- "No vaccination"
t2 <- readRDS(fls[3])
t2$vacc_type <- "Campaign"
t3 <- readRDS(fls[5])
t3$vacc_type <- "Campaign and Routine"

t <- rbind(t1, t2, t3)
t$`disability weight` <- 0.053
t$`duration of illness` <- "28 days"
t$`case fatality ratio` <- 0.01
t$`duration of vaccine-induced immunity` <- "19 years"
t$`vaccine efficacy` <- 0.95

fwrite(t, "outputs/stoch_param_Typhoid.csv")

```



## Cholera report
```{r}
devtools::load_all(".")
target_countries <- c("DRC", "Kenya", "Somalia", "South Sudan", "Ethiopia" )
cntry <- clean_country_names(target_countries)
library(data.table)

vc_list_cmp <- lapply(cntry, function(x) create_vaccine_coverage(country = x, disease = "Cholera", routine = F))

pop <- setup_cohorts(country = nms[1])
pop_vacc <- calculate_vaccine_protected(country = nms[1], population = pop, vacc_coverage = vc_list[[1]])

```

## Parameter reset
```{r}
# parameters <- fread("data/parameters.csv")
oripar <- parameters
newpar <- oripar[1:3,]
newpar$value <- newpar$value * 10
revpar <- reset_param(new = newpar, original = oripar)
```

## Output file check
```{r}
# parameters <- fread("data/parameters.csv")
library(data.table)
res <- fread("outputs/central_Typhoid_campaign.csv")
res %>% filter(country == "IND", year %in% 2023:2026) -> res1
sum(res1 < 0)
## It has negative values  
cntry <- "India"
pop <- setup_cohorts(country = cntry)
cas <- calculate_cases(disease = "Typhoid", country = cntry, population = pop)  
```
