---
title: "Typhoid final runs"
author: "Jong-Hoon Kim"
date: "3/14/2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load the package
```{r}
devtools::load_all()
```

```{r}
library(data.table)
library(devtools)
library(usethis)
library(tidyverse)
```

## Simulation using simulate_tf
### Create parameters list
```{r}
devtools::load_all()
params_list <- list()
nruns <- 200
dis <- "Typhoid"
tst <- tstamp(hour=T, minute=T, second=T)

for (k in seq_along(target_countries_tf)) {
  cntry <- target_countries_tf[[k]]
  params_list[[k]] <-
    set_params_tf(disease = dis,
                  nruns = nruns,
                  country = cntry,
                  parameter_data = parameter_data_tf,
                  incidence_rate_data = overall_incid_rate_tf)
}
#
# Added on May 14 2022 
names(params_list) <- dput(target_countries_tf)
saveRDS(params_list, paste0("outputs/params_", dis, "_", tst, ".rds"))
```

### Create parameters list - single run
Parameters including only central values for a single run
```{r}
devtools::load_all()

params_list <- list()
nruns <- 1
dis <- "Typhoid"

for (k in seq_along(target_countries_tf)) {
  cntry <- target_countries_tf[[k]]
  params_list[[k]] <-
    set_params_central_tf(disease = dis,
                  nruns = 1,
                  country = cntry,
                  parameter_data = parameter_data_tf,
                  incidence_rate_data = overall_incid_rate_tf)
}

# case_fatality_ratio is the same for <5 and 5+
for (i in seq_along(params_list)) {
  cat("i =", i, "\n")
  if (params_list[[1]]$params$case_fatality_ratio_1 !=  params_list[[1]]$params$case_fatality_ratio_2) {
    message("i = ", i)
  }
}

# Added on May 14 2022 
names(params_list) <- dput(target_countries_tf)
saveRDS(params_list, paste0("outputs/params_center_", dis, "_",               tstamp(hour=T), ".rds"))
```

create cohort list
```{r}
cohort_list <- 
  lapply(target_countries_tf, function(z) 
    setup_cohorts(country = z, year = "2000:2100", 
                  population_data = population_data))

names(cohort_list) <- target_countries_tf
saveRDS(cohort_list, paste0("outputs/cohort_list_Typhoid_",tst,".rds"))
```

### simulate_tf
```{r}
devtools::load_all()
dis <- "Typhoid"
tst1 <- "20240105T163542"
# new parameter file created including discount rate
tst2 <- "20240314T233526"
tst <- tst2 # save sim results according to the new parameter file
# params_list <- readRDS("outputs/params_center_Typhoid_20231228T14.rds")
cohort_list <- readRDS(paste0("outputs/cohort_list_Typhoid_", tst1, ".rds"))
params_list <- readRDS(paste0("outputs/params_Typhoid_", tst2, ".rds"))

run_ids <- 1:200 
# vacc_id <- 1
# for (id in seq_along(vacc_scenarios_tf)) {
cn_id0 <- 26
cn_id1 <- 45

for (vacc_id in 2:2) {
  vacc_scenario <- vacc_scenarios_tf[vacc_id]
  vc_data <- NULL
  if (vacc_id > 1) {
   vc_data <- vacc_cov_input_tf[[vacc_id - 1]] # length = 4
  }
  
  tic <- Sys.time()
  
  # disease = dis
  # year = "2000:2100"
  # country = target_countries_tf
  # params_list = params_list
  # vacc_scenario = vacc_scenario
  # vacc_cov_data = vc_data
  # cohort = cohort_list
  # run_ids = run_ids
  # 
  sim <- simulate_tf(disease = dis, 
                     year = "2000:2100",
                     country = target_countries_tf[cn_id0:cn_id1],
                     params_list = params_list,
                     vacc_scenario = vacc_scenario,
                     vacc_cov_data = vc_data,
                     cohort = cohort_list,
                     run_ids = run_ids,
                     life_expectancy_data = life_expectancy_data, 
                     save_by_country = FALSE)
  
  telapsed <- Sys.time() - tic
  cat("time elapsed = ", telapsed, "\n")
 
  # saveRDS(sim$stoch, paste0("outputs/stoch_output_list_", run_ids[1], "_", run_ids[length(run_ids)], "_", tst, "_", dis, "_", vacc_scenario, ".rds"))
  # saveRDS(sim$central, paste0("outputs/central_output_list_", run_ids[1], "_", run_ids[length(run_ids)], "_", tst, "_", dis, "_", vacc_scenario, ".rds"))
  
    saveRDS(sim$stoch, paste0("outputs/stoch_output_list_cntry_", cn_id0, "_", cn_id1, "_", tst, "_", dis, "_", vacc_scenario, ".rds"))
  saveRDS(sim$central, paste0("outputs/central_output_list_cntry_", cn_id0, "_", cn_id1, "_", tst, "_", dis, "_", vacc_scenario, ".rds"))
  
}

```



## Combine results in the list format to data.frame files

Results across countries given vaccination scenario (i.e., novacc, campaign, or
routine) are in a list type. The do.call(rbind, x) is applied to make them a 
single data.frame   

### Manual conversion
```{r}
# data in which results from each country is stored in a list
dis <- "Typhoid"
tst <- "20240314T233526"
stocen <- "central" # "stoch" #   # stoch vs central


for (vacc in vacc_scenarios_tf[1]) {
  
  res_list <- readRDS(paste0("outputs/", stocen, "_output_list_", tst, 
                             "_Typhoid_", vacc, ".rds"))
  # bind each element of the list by row
  res <- do.call(rbind, res_list)
  
  saveRDS(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tst, ".rds"))
  fwrite(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tst, ".csv"))
}

```

### Combine list sim results into data frame manually
```{r}
stocen <- "stoch" # "central" #   # stoch vs central
dis <- "Typhoid"
tst <- "20240314T233526"
vacc <- "novacc"
# res_list <- readRDS("outputs/central_output_list_1_200_20240314T233526_Typhoid_novacc.rds")
# res_list <- readRDS("outputs/stoch_output_list_1_200_20240314T233526_Typhoid_novacc.rds")
res_list <- readRDS("C:/Users/jonghoon.kim/Downloads/stoch_output_list_1_200_20240314T233526_Typhoid_novacc.rds")
res <- do.call(rbind, res_list)
# wd <- "G:/My Drive/Projects/VIMC/cholera_typhoid_vimc_ivi"
# saveRDS(res, paste0("outputs/central_", dis, "_", vacc, "_", tst, ".rds"))
# data.table::fwrite(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tst, ".csv"))
data.table::fwrite(res, paste0("C:\\Users\\jonghoon.kim\\Downloads\\VIMC_submissions\\typhoid\\", stocen, "_", dis, "_", vacc, "_", tst, ".csv"))


# data.table::fwrite(res, paste0("C:/Users/jonghoon.kim/Downloads/outputs/", stocen, "_", dis, "_", vacc, "_", tst, ".csv"))  
```

### Combine sim results by run ids

```{r}
devtools::load_all()
stocen <- "stoch"
tst <- "20240105T163542"
vacc <- "campaign_routine_bluesky"
dis <- "Typhoid"
# stoch_output_list_151_180_20240105T163542_Typhoid_campaign_bluesky.rds
fls <- list.files("outputs", pattern = paste0(".*", stocen, ".*output.*list.*", tst, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)

reslist <- vector('list', length(fls))
for (i in seq_along(fls)) {
  reslist[[i]] <- do.call(rbind, readRDS(fls[i]))
  # reslist[[i]] <- data.table::rbindlist(readRDS(fls[i]))
}

# purrr::map_chr(reslist, nrow)
res <- data.table::rbindlist(reslist)

# saveRDS(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tst, ".rds"))
data.table::fwrite(res, 
  paste0("C:\\Users\\jonghoon.kim\\Downloads\\VIMC_submissions\\typhoid\\",
    stocen, "_", dis, "_", vacc, "_", tst, ".csv"))
# res2 <- data.table::fread(paste0("C:\\Users\\jonghoon.kim\\Downloads\\VIMC_submissions\\typhoid\\stoch_Typhoid_novacc_20240105T163542.csv"))
```

### Combine sim results by countries

```{r}
devtools::load_all()
stocen <- "stoch"
tst <- "20240314T233526"
dis <- "Typhoid"
for (x in 5:5){
vacc <- vacc_scenarios_tf[x]
vacc <- "campaign_routine_bluesky"
# stoch_output_list_151_180_20240105T163542_Typhoid_campaign_bluesky.rds
fls <- list.files("outputs", pattern = paste0(".*", stocen, ".*output.*list.*",
    tst, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)

reslist <- vector('list', length(fls))
for (i in seq_along(fls)) {
  reslist[[i]] <- do.call(rbind, readRDS(fls[i]))
  # reslist[[i]] <- data.table::rbindlist(readRDS(fls[i]))
}

# purrr::map_chr(reslist, nrow)
res <- data.table::rbindlist(reslist)

# saveRDS(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tst, ".rds"))
data.table::fwrite(res, 
  paste0("C:\\Users\\jonghoon.kim\\Downloads\\VIMC_submissions\\typhoid\\",
    stocen, "_", dis, "_", vacc, "_", tst, ".csv"))
# res2 <- data.table::fread(paste0("C:\\Users\\jonghoon.kim\\Downloads\\VIMC_submissions\\typhoid\\stoch_Typhoid_novacc_20240105T163542.csv"))
}
```


## Output diagnostics  

Compare w/ and w/o indirect effects
```{r}
sim1 <- readRDS("outputs/central_Typhoid_routine_1_20211125.rds")
sim2 <- readRDS("outputs/central_Typhoid_routine_1_20211217T1117.rds")

# cntry = target_countries_tf[1]
dis = "Typhoid"

library(ggplot2)
library(gridExtra)
for (cntry in target_countries_tf){
  plt1 <- year_val_plot(disease = dis, country = cntry, 
                        sim_central_value = sim1)
  plt2 <- year_val_plot(disease = dis, country = cntry, 
                        sim_central_value = sim2)
  grid.arrange(plt1, plt2, nrow=2)
  invisible(readline(prompt = "Press RET to continue"))
}
```


### Diagnostic plots
For countries like Guyana, Haiti, why DALYs outputs show step-like lines? 
1) it appears that the step occurs every 5 years

```{r}
library(ggplot2)
# tst <- "20240102"
dis <- "Typhoid"
for (cntry in target_countries_tf) {
  # vacc <- "campaign_bluesky"
  # vacc <- "campaign_routine_bluesky"
  # vacc <- "campaign_default"
  # vacc <- "campaign_routine_default"
  # vacc <- "novacc"
  # cntry <- "Iran, Islamic Republic of"
  # res <- readRDS(paste0("outputs/central_", dis, "_", vacc, "_", tst, ".rds"))
  plt <- year_val_plot(disease = dis, country = cntry, sim_central_value = res)
  # year_val_plot(disease = dis, country = cntry, sim_central_value = res, age = 10)
  print(plt)
  invisible(readline(prompt = "Press RETURN to continue"))
} 
```

### Diagnostics I
DALYs may have step increase or decrease every 5 years because life expectancy
is given every 5 years and I assume that  

```{r}
dis <- "Typhoid"
# Dropbox <- "C:\\Users\\jonghoon.kim\\Dropbox\\VIMC\\submitted_12May2022\\"
path_to_folder <- "G:\\My Drive\\Projects\\VIMC\\cholera_typhoid_vimc_ivi\\outputs\\"

vacc <- "novacc"
tst <- "20240104T18"
nv <- readRDS(paste0(path_to_folder, "central_", dis, "_", vacc, "_", tst, ".rds"))

vacc <- "campaign_default"
c1 <- readRDS(paste0(path_to_folder, "central_", dis, "_", vacc, "_", tst, ".rds"))
vacc <- "campaign_routine_default"
r1 <- readRDS(paste0(path_to_folder, "central_", dis, "_", vacc, "_", tst, ".rds"))
vacc <- "campaign_bluesky"
c2 <- readRDS(paste0(path_to_folder, "central_", dis, "_", vacc, "_", tst, ".rds"))
vacc <- "campaign_routine_bluesky"
r2 <- readRDS(paste0("outputs/central_", dis, "_", vacc, "_", tst, ".rds"))

datfull <- list()
datfull[[1]] <- nv
datfull[[2]] <- r1
datfull[[3]] <- r2
datfull[[4]] <- c1
datfull[[5]] <- c2

datsub <- list()

cntries <- unique(r1$country_name)
for(cn in cntries){
  cat(paste0("country = ", cn))
  ca_nv <- round(sum(nv[nv$country_name == cn,]$cases))
  ca_r1 <- round(sum(r1[r1$country_name == cn,]$cases))
  ca_r2 <- round(sum(r2[r2$country_name == cn,]$cases))
  ca_c1 <- round(sum(c1[c1$country_name == cn,]$cases))
  ca_c2 <- round(sum(c2[c2$country_name == cn,]$cases))
  cat(paste0(": ", ca_nv, ", ", ca_c1, ", ", ca_c2, ", ", ca_r1, ", ", ca_r2,"\n"))
  cat(paste0("Diff: ", ca_nv - ca_c1, ", ", ca_nv - ca_c2, ", ", ca_nv - ca_r1, ", ", ca_nv - ca_r2, "\n"))
  invisible(readline(prompt = "Press RETURN to continue"))
}
```


```{r}
# plot against time
vacc_scen <- c("novacc", "campaign_default", "campaign_routine_default", "campaign_bluesky", "campaign_routine_bluesky")

library(ggplot2)

for(cn in cntries) {
  datsub <- list()
  for(i in 1:length(datfull)) {
    datfull[[i]] %>%
      filter(country_name == cn) %>%
      group_by(year) %>%
      summarize(cohort_size = sum(cohort_size),
                cases = sum(cases),
                deaths = sum(deaths),
                dalys = sum(dalys),
                yll = sum(yll),
                vacc_scenario = vacc_scen[i]) -> dtemp
    
      dtemp$vacc_scenario = vacc_scen[i] 
      datsub[[i]] <- dtemp
  }
  df <- do.call(rbind, datsub)
  plt <- ggplot(df)+
    geom_line(aes(year, cases, color = vacc_scenario))+
    scale_colour_manual(values = c("darkred", "red", "black", "blue",
                                   
                                   "steelblue"))+
    ggtitle(paste0(cn))
  
  print(plt)
  
  invisible(readline(prompt = "Press RETURN to continue"))
}

```


### Diagnostic plots II
Visualize cohort size, cases, deaths, and DALYs across vaccination scenarios
for each of the countries over the period of 2000 - 2100

```{r}
library(ggplot2)
library(grid)
library(gridExtra)
dis <- "Typhoid"
tst <- "20240104T18"
vacc_scenarios <- c("campaign_default", "campaign_routine_default", 
                    "campaign_bluesky", "campaign_routine_bluesky")
vacc <- vacc_scenarios[2]

for (cntry in target_countries_tf){
  res1 <- readRDS(paste0("outputs/central_",  dis, "_", "novacc_", tst, ".rds"))
  res2 <- readRDS(paste0("outputs/central_",  dis, "_", vacc, "_", tst, ".rds"))
  
  p1 <- year_val_plot(disease = dis, country = cntry, sim_central_value = res1, vacc_scenario = "novacc")
  p2 <- year_val_plot(disease = dis, country = cntry, sim_central_value = res2, vacc_scenario = "campaign_default")

  plt <- grid.arrange(p1, p2, ncol = 1)
  print(plt)
  invisible(readline(prompt = "Press RETURN to continue"))
} 
```

### Diagnostic III
1. Check if there are population categories (defined by chronological year
and age) have negative numbers
2. Check if cohort sizes are same across vaccination scenarios 
3. Check if cases, deaths, and DALYs are larger for non-vaccination scenarios
compared to vaccination scenarios
4. Check if cases, deaths, and DALYs are larger for routine (or campaign)
scenario compared to the non-vaccination scenario

```{r}
dis <- "Typhoid"
# tst <- "20240104T13"
tst <- "20240102"
vacc_scenarios <- c("campaign_default", "campaign_routine_default",
                    "campaign_bluesky", "campaign_routine_bluesky")

for (vacc in vacc_scenarios) {
  cat("vacc =", vacc, "\n")
  res1 <- readRDS(paste0("outputs/central_",  dis, "_", "novacc_", tst, ".rds"))
  res2 <- readRDS(paste0("outputs/central_",  dis, "_", vacc, "_",
                         tst, ".rds"))
  
  if (sum(res1$cohort_size != res2$cohort_size) != 0) {
    message("Cohort size of novacc and vacc scenrios are different for", 
            cntry, "\n")
  }
  if (sum(res1$cases < res2$cases) != 0) {
    message("Cases are bigger for vacc scenario than novacc scenrios for", 
            cntry, "\n")    
  }
   if (sum(res1$deaths < res2$deaths) != 0) {
    message("Deaths are bigger for vacc scenario than novacc scenrios for", 
            cntry, "\n")    
  }
  if (sum(res1$dalys < res2$dalys) != 0) {
    message("DALYs are bigger for vacc scenario than novacc scenrios for", 
            cntry, "\n")
  }
  if (sum(res1$yll < res2$yll) != 0) {
    message("YLLs are bigger for vacc scenario than novacc scenrios for",
            cntry, "\n")
  }
}
```

## Montagu-compatible csv files

### Prepare csv files for uploading the results onto Montagu
```{r}
library(data.table)
dis <- "Typhoid"
tst <- "20240104T18"

for (vacc in c("novacc", "campaign_default", "campaign_routine_default",
               "campaign_bluesky", "campaign_routine_bluesky")) {
  cat("vacc =", vacc, "\n")
  
  fl <- list.files("outputs/", 
                   pattern = paste0("central.*", dis, ".*", vacc,
                                                ".*", tst,  ".*rds$"),
                   full.names = T)
  d <- readRDS(fl)
  fwrite(d, paste0("outputs/vimc_central_", dis, "_", vacc, "_", tst, ".csv"))
}
```

### Prepare csv files for the parameters
```{r}
tst <- "20240314T233526"
params_list <- readRDS(paste0("outputs/params_Typhoid_", tst, ".rds"))

## Combine parameters
lst <- list()
for (i in seq_along(target_countries_tf)) {
  params <- params_list[[i]]
  lst[[i]] <- params$params
  # ir <- params[[1]]
  # t_ir <- data.table::transpose(ir)
  # colnames(t_ir) <- paste0("incid_rate_ag_", 0:100)
  # t_cfr <- data.table::transpose(params$case_fatality_ratio)
  # colnames(t_cfr) <- paste0("case_fatality_ratio_ag_", 0:100)
  # lst[[i]] <- cbind(params$params, t_cfr)
}

res <- do.call(rbind, lst)
# have separate values dependent on age and 
# were not used for simulation
# res$case_fatality_ratio <- NULL 
# res <- do.call(rbind, params_list)
dis <- "Typhoid"
# consistent with the name of the simulation results
# saveRDS(res, paste0("outputs/params_", dis, "_", tst, ".rds"))
# fwrite(res, paste0("outputs/vimc_params_", dis, "_", tst, ".csv"))

data.table::fwrite(res, 
  paste0("C:\\Users\\jonghoon.kim\\Downloads\\VIMC_submissions\\typhoid\\",
    "stoch_parameters_", dis, "_", tst, ".csv"))
```


# Diagnostic IV: Negative population size

Negative values can happen when there is population size for a certain age group
is zero and in the previous year there are vaccinated people in one-year younger
age group some value and model predicts some people move to the next age group,
which is determined by the probability of dying and aging, vaccine  

```{r}
vacc_scenario <- c("novacc", "campaign_default", "campaign_routine_default",
                   "campaign_bluesky", "campaign_routine_bluesky")

tst <- "20240104T18"

for (vacc in vacc_scenario) {
  fl <- paste0("outputs/vimc_central_Typhoid_", vacc, "_", tst, ".csv")
  message(paste0("file = ", fl))
  d <- fread(fl)
  if(sum(d$cohort_size < 0) > 0){
    message("Negative cohort size!")
  }
  if(sum(d$cases < 0) > 0){
    message("Negative cases!")
  }
  if(sum(d$deaths < 0) > 0){
    message("Negative deaths!")
  }
  if(sum(d$dalys < 0) > 0){
    message("Negative DALYs!")
  }
  if(sum(d$yll < 0) > 0){
    message("Negative YLLs!")
  }
}
```


## Check cohort size
```{r}
library(data.table)
dis <- "Typhoid"
tst <- "20240104T18"

for (vacc in c("novacc", "campaign_default", "campaign_routine_default",
               "campaign_bluesky", "campaign_routine_bluesky")) {
  
  cat("vacc =", vacc, "\n")
  res0 <- fread(paste0("outputs/vimc_central_", dis, "_novacc_", tst, ".csv"))
  res1 <- fread(paste0("outputs/vimc_central_", dis, "_", vacc, "_", tst, ".csv"))
  message(sum(res0$cohort_size != res1$cohort_size))
  message(sum(res0$cohort_size - res1$cohort_size))
}

```

## Checklist
1. Download the data set and familiarize yourself with the system
2. Check if the each function works for each data set provided
  a. Country names may differ between different data sets
  b. See if there are any missing values
## Set-up
```{r setup}
# country names from the report template
# template <- 
#   fread("inst/extdata/central-burden-template.202110gavi-3.Typhoid_IVI-Kim_standard.csv")
template <- read_csv("inst/extdata/central-burden-template.202310gavi-4.Typhoid_IVI-Kim_standard.csv")
countries <- clean_country_names(unique(template$country_name))
dis <- "typhoid"
year_chr <- "2000:2100"
```

## Steps
```{r, include=FALSE}
cntry <- countries[13]

for (cntry in countries) {
  pop <- setup_cohorts(country = cntry, 
                       year = eval(parse(text = year_chr)),
                       population_data = population_data)
}

# vacc_cov_data <- list()
# vacc_cov_data[[1]] <- vacc_cov_input_routine_tf
# vacc_cov_data[[2]] <- vacc_cov_input_campaign_tf
# vacc_cov_data[[3]] <- vacc_cov_input_routine_ia2030_tf
# vacc_cov_data[[4]] <- vacc_cov_input_campaign_ia2030_tf

# i <- 1
# vacc_cov_data[[i]] %>%
#   filter(country_code == "CAF", year == 2024) -> vd
  
# vacc_cov_input_tf from 'typhoid_final_data_prep_2023.Rmd' file
vacc_cov_data <- vacc_cov_input_tf 

pop <- as.data.frame(pop)
for (i in seq_along(vacc_cov_data)) {
  for (cntry in countries) {
    cat("vacc_cov_data_i =", i, ", country =", cntry, "\n")
    vc <- create_vaccine_coverage(country = cntry,
                            population = pop,
                            vaccine_coverage_data = vacc_cov_data[[i]])
    
    vacc_protected_direct <- 
      calculate_vaccine_protected(disease = dis,
                                  country = cntry,
                                  population = pop,
                                  year = 2000:2100,
                                  vaccine_coverage = vc,
                                  vaccine_efficacy = 0.84,
                                  vaccine_immunity_duration = 15,
                                  exponential_decay = TRUE,
                                  vaccine_efficacy_by_year = NULL)
  
  }
}
# population is now 
pop_unprotected <- pop - vacc_protected_direct
sum(pop_unprotected < 0)

# ve <- c(0.816, 0.588,0.918)
# cov <- 0.64*30685/155448
# eff_cov <- ve*cov

# indirect vaccine protection
# first calculate 

# vacc_recipients <- vc * pop
# vaccine recipients decrease only by death and for convenience vaccine immunity 
# duration is set to be very large 

#----------------------------------
# Added on 29 April 2022
vacc_protected <- vacc_protected_direct
#---------------------------------------

eff_vacc_cov_year <- calc_vacc_cov_year(pop = pop, vacc_recipients = vacc_protected)

vacc_recipients <- 
      calculate_vaccine_protected(disease = dis,
                                  country = cntry,
                                  population = pop,
                                  year = 2000:2100,
                                  vaccine_coverage = vc,
                                  vaccine_efficacy = 1.0,
                                  vaccine_immunity_duration = 1000,
                                  exponential_decay = TRUE,
                                  vaccine_efficacy_by_year = NULL)
pop_unvacc <- pop - vacc_recipients
indirect_vacc_protected <- 
  calc_indirect_vacc_protected(disease="Typhoid", country=cntry,
                               population=pop_unvacc,
                               vacc_cov=eff_vacc_cov_year,
                               vacc_cov_limit = 0.05, indirect_vacc_efficacy = 0.2)

pop_unprotected <- pop_unprotected - indirect_vacc_protected
sum(pop_unprotected < 0)


# Impact of vaccine waning
vacc_cov_year <- calc_vacc_cov_year(pop = pop, vacc_recipients = vacc_recipients)
eff_vacc_cov_year <- calc_vacc_cov_year(pop = pop, vacc_recipients = vacc_protected)

plot(1:101, vacc_cov_year, type = "l", xlab = "Year since 2000",
     main = "Vaccine recipients who lost immunity\ndue to waning and imperfect protection")
lines(1:101, eff_vacc_cov_year, col = 2)                              


# # sample an incidence rate
# overall_ir_dat %>%
#   filter(country == cntry) %>%
#   select(min, max) %>% unlist() -> ir_val
# 
# set.seed(1)
# overall_ir <- runif(1, min = ir_val[1], max = ir_val[2])

overall_ir <- 150

age_spec_ir_ref <- 
  calc_age_spec_ir_ref_tf(country = cntry, 
                         incid = incid_tf_ihme,
                         ref_year = 2017)
## popultion = population_data removed on 2023-12-26 
# ir_adj <- calc_incid_rate_risk_adj(country = cntry,
#                         overall_ir = overall_ir,
#                         wash_risk_ratio = wash_risk_ratio,
#                         wash_prop = wash_prop, 
#                         ref_year = 2017,
#                         population = population_data,
#                         incid = incid_tf_ihme)

ir_adj <- calc_incid_rate_risk_adj(country = cntry,
                        overall_ir = overall_ir,
                        wash_risk_ratio = wash_risk_ratio,
                        wash_prop = wash_prop, 
                        ref_year = 2017,
                        incid = incid_tf_ihme)

# check if the incidence rates pass through the reference
# age_specific incidence rate
age_ir_ref <- overall_ir * age_spec_ir_ref$ir_dist
plot(rep(17,101), age_ir_ref, xlim = c(1,101))
for (i in 1:101) {
  lines(1:101, ir_adj[i,], col=2)
}

cases <- calculate_cases_tf(country = cntry,
                             population = pop_unprotected,
                             year = 2000:2100,
                             incidence_rate = ir_adj)

pop[1:10,1]
cases[1:10,1]
head(pop)
head(cases)

pars_data <- parameters

dis_sev <- extract_param_tf(params_data = pars_data, str = str_dis_severity)
dis_dur <- extract_param_tf(params_data = pars_data, str = str_dis_duration)
dis_wt <- extract_param_tf(params_data = pars_data, str = str_dis_disability)

yld <- calculate_YLD_tf(cases = cases,
                        disability_weight = dis_wt,
                        prob_severity = dis_sev,
                        illness_duration = dis_dur)
yld[1:10,1:20]

# ir_data$val[1:10]* sum(dis_sev * dis_wt * dis_dur)
# yld_data$val[1:10]


# This needs to be wrapped in a function
# Case fatality ratio by age
# Data extracted from the global average (Stanaway et al. 2019 -- GBD 2017)
ir_age <- fread("inst/extdata/ir_age_typhoid.csv")
mort_age <- fread("inst/extdata/mort_age_typhoid.csv")

ir_lower <- ir_age %>% filter(name == "lower")
ir_upper <- ir_age %>% filter(name == "upper")

nrow(ir_lower) == nrow(ir_upper)
nruns <- 200
ir_sample <- data.frame(matrix(NA, nrow(ir_lower), ncol = nruns))
for (i in 1:200) {
  ir_sample[,i] <- runif(nrow(ir_lower), min = ir_lower$value, max = ir_upper$value)
}

mort_lower <- mort_age %>% filter(name == "lower")
mort_upper <- mort_age %>% filter(name == "upper")

nrow(mort_lower) == nrow(mort_upper)
nruns <- 200
mort_sample <- data.frame(matrix(NA, nrow(mort_lower), ncol = nruns))
for (i in 1:200) {
  mort_sample[,i] <- runif(nrow(mort_lower), min = mort_lower$value, max = mort_upper$value)
}

cfr_sample <- mort_sample / ir_sample
apply(cfr_sample, 1, summary)  
apply(cfr_sample, 1, mean) 

cfr <- calc_cfr(country = cntry)

yll <- calculate_YLL_tf(cases = cases,
                        country = cntry,
                        case_fatality = cfr)


yll[1:10,"2015"]
cases[1:10, "2015"] * cfr[1:10] * life_expectancy[1:10, "2015"] # 0-9 y0

ir[1:10] * cfr[1:10] * life_expectancy[1:10, "2015"] # 0 -9 y0
yll_data[yll_data$age == "1 to 4",]$val
yll_data[yll_data$age == "5 to 9",]$val
yll_data[yll_data$age == "<1 year",]$val
# it appears that life expectancy is lower for us compared to the IHME

daly <- calculate_DALY_tf(country = cntry,
                          cases = cases,
                          disability_weight = dis_wt,
                          prob_severity = dis_sev,
                          illness_duration = dis_dur, 
                          case_fatality = cfr)
daly[1:10, 1:10]
yll[1:10, 1:10] + yld[1:10, 1:10]
```



## Sampling parameters
```{r}
# Sampling based on Sobol low discrepancy sequence   
countries
nruns <- 5
parameter_data <- parameters
life_expectancy_data <- life_expectancy
# edited on 2023-12-26
# vacc_cov_data <- vacc_cov_input_routine_tf
vacc_cov_data <- vacc_cov_input_tf[[1]]

countries
cntry <- countries[1]
yr <- 2000:2100
dis <- "Typhoid"
params_list <- 
  set_params_tf(disease = dis,
                nruns = nruns,
                country = cntry,
                parameter_data = parameter_data,
                incidence_rate_data = overall_incid_rate)

# get country-level fixed parameters
life_exp <- 
  get_life_expectancy(life_expectancy_data = life_expectancy_data, country = cntry, year = yr)

vi <- vaccine_impact_tf(disease = dis,
                  country = cntry,
                  population_data = population_data,
                  vaccine_coverage_data = vacc_cov_data,
                  vacc_scenario = "routine",
                  year = yr,
                  life_expectancy = life_exp,
                  # case_fatality_ratio = params_list$case_fatality_ratio[,1],
                  parameter_sample = params_list$params[1,]) 

```

# Comparison of vaccination programs
```{r}
library(data.table)
library(dplyr)
# routine_tf_gavi2  <- fread("inst/extdata/coverage_202110gavi-2_typhoid-routine-default.csv")
# routine_tf_gavi3  <- fread("inst/extdata/coverage_202110gavi-3_typhoid-routine-default.csv")
# campaign_tf_gavi2 <- fread("inst/extdata/coverage_202110gavi-2_typhoid-campaign-default.csv")
# campaign_ia2030_tf_gavi3 <- fread("inst/extdata/coverage_202110gavi-3_typhoid-campaign-ia2030_target.csv")
# routine_ia2030_tf_old <- fread("inst/extdata/coverage_202110gavi-2_typhoid-routine-ia2030_target.csv")
# 

campaign_default_tf <- fread("inst/extdata/coverage_202310gavi-4_typhoid-campaign-default.csv")
campaign_routine_default_tf <- fread("inst/extdata/coverage_202310gavi-4_typhoid-campaign-routine-default.csv")
campaign_bluesky_tf <- fread("inst/extdata/coverage_202310gavi-4_typhoid-campaign-bluesky.csv")
campaign_routine_bluesky_tf <- fread("inst/extdata/coverage_202310gavi-4_typhoid-campaign-routine-bluesky.csv")

all.equal(campaign_bluesky_tf, campaign_default_tf) 
cntries <- unique(campaign_bluesky_tf$country)

for (cntry in cntries) {
  cat(cntry, "\n")
  campaign_default_tf %>% filter(country == cntry) -> default
  campaign_bluesky_tf %>% filter(country == cntry) -> bluesky
  if(!isTRUE(all.equal(default, bluesky))){
    print(default)
    print(bluesky)
  }
}

for (cntry in cntries) {
  campaign_default_tf %>% filter(country == cntry) -> default
  if(nrow(default) > 1){
    cat(cntry, "\n")
  } 
}

# The following 5 countries include two vaccination campaigns in the default
# campaign whereas the bluesky scenario includes only one vaccination 
# campaign in earlier years.

# Congo, the Democratic Republic of the 
# Ethiopia 
# Indonesia 
# India 
# Nigeria 
# Pakistan 
# Tanzania, United Republic of 

campaign_default_tf %>% filter(country == "Pakistan")
campaign_bluesky_tf %>% filter(country == "Pakistan")
        
```

```{r}
countries <- clean_country_names(unique(template$country_name))

yr <- 2000:2100

vacc_cov_input <- list()
vacc_cov_input[[1]] <- vacc_cov_input_campaign_default_tf
vacc_cov_input[[2]] <- vacc_cov_input_campaign_routine_default_tf
vacc_cov_input[[3]] <- vacc_cov_input_campaign_bluesky_tf
vacc_cov_input[[4]] <- vacc_cov_input_campaign_routine_bluesky_tf

# for(r in 1:length(countries)){
for(r in 60:60){
  cntry <- countries[r]
  message(paste0("country = ", cntry))
  
  pop <- setup_cohorts(country = cntry,
                       year = yr,
                       population_data = population_data)
  
  pop <- as.data.frame(pop)
  for(i in 4:4) {
    vacc_cov <- create_vaccine_coverage(country = cntry,
                                population = pop,
                                vaccine_coverage_data = vacc_cov_input[[i]])
  }
}
  
# invisible(readline(prompt = "Press RET to continue"))

```

## Parameter transformation test
```{r}
# incidence rate
incidence_rate_data %>% filter(country == cntry) %>%
  pull(max) -> ir_ub
incidence_rate_data %>% filter(country == cntry) %>%
  pull(min) -> ir_lb

# duration of illness in years
disdur_lb <- extract_param_tf(parameter_data, str_dis_duration, "lower")
disdur_ub <- extract_param_tf(parameter_data, str_dis_duration, "upper")
disdur_center <- extract_param_tf(parameter_data, str_dis_duration, "val")
# disability weight
diswt_lb <- extract_param_tf(parameter_data, str_dis_disability, "lower")
diswt_ub <- extract_param_tf(parameter_data, str_dis_disability, "upper")
diswt_center <- extract_param_tf(parameter_data, str_dis_disability, "val")
# no uncertainty range are are available for prob of disease severity
disseverity_val <- extract_param_tf(parameter_data, str_dis_severity, "val")

str <- "duration of vaccine-induced immunity (year)"
vaccdur_lb <- extract_param_tf(parameter_data, str, "lower")
vaccdur_ub <- extract_param_tf(parameter_data, str, "upper")
vaccdur_center <- extract_param_tf(parameter_data, str, "val")

vacceff_lb <- extract_param_tf(parameter_data, "vaccine efficacy", "lower")
vacceff_ub <- extract_param_tf(parameter_data, "vaccine efficacy", "upper")
vacceff_center <- extract_param_tf(parameter_data, "vaccine efficacy", "val")

  
set.seed(42)
cntry <- countries[[1]]
# uniform random parameters are transformed within the function
params <- 
  set_params_tf(disease = "Typhoid",
                nruns = 2000,
                country = cntry,
                parameter_data = parameter_data,
                incidence_rate_data = overall_incid_rate_tf)
i=4  
for (i in 1:4){
  plot(density(eval(parse(text=paste0("params$params$disability_weight_",i)))))
  abline(v=c(diswt_lb[i], diswt_ub[i]), col=c(2,2))
  
  plot(density(eval(parse(text=paste0("params$params$illness_duration_",i)))))
  abline(v=c(disdur_lb[i], disdur_ub[i]), col=c(2,2))
}

plot(density(eval(parse(text=paste0("params$params$vacc_efficacy")))))
abline(v=c(vacceff_lb, vacceff_ub), col=c(2,2))

plot(density(eval(parse(text=paste0("params$params$duration_vacc_protection")))))
abline(v=c(vaccdur_lb, vaccdur_ub), col=c(2,2))

plot(density(eval(parse(text=paste0("params$params$incidence_rate")))))
abline(v=c(ir_lb, ir_ub), col=c(2,2))

```

## Vaccine coverage over time
```{r}
template <- 
  fread("inst/extdata/central-burden-template.202110gavi-3.Typhoid_IVI-Kim_standard.csv")
countries <- unique(template$country_name)
dis <- "Typhoid"
year_chr <- "2000:2100"
cntry <- countries[1]

for (cntry in countries) {
  pop <- setup_cohorts(country = cntry, 
                       year = eval(parse(text = year_chr)),
                       population_data = population_data)
}

vacc_cov_data <- list()
vacc_cov_data[[1]] <- vacc_cov_input_routine_tf
vacc_cov_data[[2]] <- vacc_cov_input_campaign_tf
vacc_cov_data[[3]] <- vacc_cov_input_routine_ia2030_tf
vacc_cov_data[[4]] <- vacc_cov_input_campaign_ia2030_tf
i=1
for (i in seq_along(vacc_cov_data)) {
  for (cntry in countries) {
    cat("vacc_cov_data_i =", i, ", country =", cntry, "\n")
    vc <- create_vaccine_coverage(country = cntry,
                            population = pop,
                            vaccine_coverage_data = vacc_cov_data[[i]])
    
    vacc_protected <- 
      calculate_vaccine_protected(disease = dis,
                                  country = cntry,
                                  population = pop,
                                  year = 2000:2100,
                                  vaccine_coverage = vc,
                                  vaccine_efficacy = 0.81,
                                  vaccine_immunity_duration = 15,
                                  exponential_decay = TRUE,
                                  vaccine_efficacy_by_year = NULL)
  
  }
}
## population is now 
pop_unvacc <- pop - vacc_protected
sum(pop_unvacc < 0)

pop_yr <- apply(pop, 2, sum)
vacc_protected_yr <- apply(vacc_protected, 2, sum)

# calc_vacc_cov_year <- function(pop, vacc_procted){
#   pop_yr <- apply(pop, 2, sum)
#   vacc_protected_yr <- apply(vacc_protected, 2, sum)
#   # prop_vacc_yr <- vacc_protected_yr / pop_yr
#   return(vacc_protected_yr / pop_yr)
# }

vacc_cov_year <- calc_vacc_cov_year(pop, vacc_procted)

prop_vacc_yr <- vacc_protected_yr / pop_yr
pop_yr_u15 <- apply(pop[1:15,], 2, sum)
vp_yr_u15 <- apply(vacc_protected[1:15,], 2, sum)
prop_vacc_yr_u15 <- vp_yr_u15 / pop_yr_u15

plot(2000:2100, prop_vacc_yr, type="l", ylim=c(0,1))
lines(2000:2100, prop_vacc_yr_u15, type="l", col=2)


# Bangladesh cluster-randomzied trial
#an average vaccine coverage of children aged 9 months to less than 16 years
#of 64% during the follow-up period, Qadri et al. (2021) Lancet
cov_overall <- 0.64*30685/155448
ive <- 0.19
prop_vacc_yr
# Overall vaccine coverage was around 0.1263342*100 and the indirect VE was
# 19% (-12, 41)
# We made a simple  
# This is a somewhat arbitrary decision and we explore 10% coverage and we explore various levels of vaccine coverage

vacc_cov_year <- calc_vacc_cov_year(pop, vacc_procted)
indirect_vacc_protected <- 
  calc_indirect_vacc_protected(disease="Typhoid", country=cntry,
                               population=pop_unvacc,
                               vacc_cov=vacc_cov_year,
                               vacc_cov_limit = 0.1, indirect_vacc_efficacy = 0.2)

for(i in 1:101) {
  plot(1:101, vacc_cov_year, type="l",
       xlab="Year since 2000", main=paste0("Age = ", i-1, ", VC, IVP/UV"))
  abline(h=0.1, col=2)
  lines(1:101, indirect_vacc_protected[i,]/pop_unvacc[i,], col=3, type="l")
  readline(prompt = "Press RET to continue")
}
# by year
for(i in 1:nrow(vacc_protected)){
  cat("i=", i, "\n")
  plot(1:101, pop[i,], type="l", ylim=c(0,max(pop)), 
       main=paste0("Age = ", i-1, ", Pop, VP, IVP"))
  lines(1:101, vacc_protected[i,], col=2)
  lines(1:101, indirect_vacc_protected[i,], col=3)
  readline(prompt = "Press RET to continue")
}

for(i in 1:nrow(vacc_protected)){
  cat("i=", i, "\n")
  plot(1:101, vacc_protected[i,]/pop[i,], type="l", ylim=c(0, max(vacc_protected/pop)), main=paste0("Age = ", i-1, ", VP/Pop, IVP/Pop"))
  lines(1:101, indirect_vacc_protected[i,]/pop[i,], col=2)
  readline(prompt = "Press RET to continue")
}

i=29
# by age
for(i in 1:ncol(vacc_protected)){
  cat("i=", i, "\n")
  plot(1:101, vacc_protected[,i], ylim=c(0, max(pop[,i])), type="l", main="By year")
  lines(1:101, indirect_vacc_protected[,i], col=2, type="l")
  readline(prompt = "Press RET to continue")
}

```

## Simulation
Some methods require disease and other don't. How do I streamline it?
```{r}
# Restart R (Ctrl + Shift + F10) to remove any remaining objects 
#
devtools::load_all()
set.seed(42)
nruns <- 30
#VIMC report requires capital T in Typhoid
dis <- "Typhoid"
year <- "2000:2100"
vaccination_scenarios <- c("novacc", "routine", "routine", "campaign", "campaign")
vaccination_scenarios_save <- c("novacc", "routine_1", "routine_2", "campaign_1", "campaign_2")

tstamp <- format(Sys.time(), "%Y%m%dT%H%M")
# target countries
template <- 
  fread("inst/extdata/central-burden-template.202110gavi-3.Typhoid_IVI-Kim_standard.csv")
countries <- clean_country_names(unique(template$country_name))
# parameters to explore
# library(magrittr)
parameter_data <- parameters %>% filter(tolower(disease) == tolower(dis))

# # Sampling based on Sobol low discrepancy sequence
# params_list <- list()
# for (k in 1:length(countries)) {
#   cntry <- countries[[k]]
#   params_list[[k]] <-
#     set_params_tf(disease = dis,
#                   nruns = nruns,
#                   country = cntry,
#                   parameter_data = parameter_data,
#                   incidence_rate_data = overall_incid_rate_tf)
# }
# 
# saveRDS(params_list, paste0("outputs/params_", dis, "_", tstamp, ".rds"))

# params_list <- readRDS("outputs/params_Typhoid_20211217T1101.rds")
params_list <- readRDS("outputs/params_Typhoid_20220509T2324.rds")
stoch_output <- list()
stoch_parameters <- list()
central_output <- list()

# four vaccine coverage inputs
# vacc_cov_input_routine_tf
# vacc_cov_input_campaign_ia2030_tf
# vacc_cov_input_campaign_tf
# vacc_cov_input_routine_ia2030_tf

vacc_cov_input <- list()
vacc_cov_input[[1]] <- vacc_cov_input_routine_tf
vacc_cov_input[[2]] <- vacc_cov_input_routine_ia2030_tf
vacc_cov_input[[3]] <- vacc_cov_input_campaign_tf
vacc_cov_input[[4]] <- vacc_cov_input_campaign_ia2030_tf

## arrange vacc_scenarios according to the vaccination scenarios

vacc_cov_data <- vacc_cov_input[[1]]

# for (r in seq_along(vaccination_scenarios)) {
for (r in 1:1) {
  vacc_scenario <- vaccination_scenarios[r]
  vacc_scenario_save <- vaccination_scenarios_save[r]
  if (r > 1) {
    vacc_cov_data <- vacc_cov_input[[r - 1]]
  }
  for (k in seq_along(countries)) {
    cntry <- countries[k]
    cohort <- list()
    cases <- list()
    deaths <- list()
    dalys <- list() 

    life_exp <- 
      get_life_expectancy(life_expectancy_data = life_expectancy,
                          country = cntry,
                          year = eval(parse(text = year)))
    
    message(paste0("vaccine scenario = ", vacc_scenario, ", country = ", cntry))
    ## country-specific parameters
    params <- params_list[[k]] # country by index
    
    for (i in 1:nruns) {
      # message(paste0("r = ", r, ", k = ", k, ", i = ", i))
      param_sample <- list()
      param_sample$vacc_efficacy <- params$params$vacc_efficacy[i]
      param_sample$duration_vacc_protection <- 
        params$params$duration_vacc_protection[i]
      param_sample$indirect_vacc_efficacy <-
        params$params$indirect_vacc_efficacy[i]
      param_sample$eff_vacc_cov_limit <- 
        params$params$eff_vacc_cov_limit[i]
      
      param_sample$illness_duration <- 
        c(params$params$illness_duration_1[i],
          params$params$illness_duration_2[i],
          params$params$illness_duration_3[i],
          params$params$illness_duration_4[i])

      param_sample$prob_severity <- 
        c(params$params$prob_severity_1[i],
          params$params$prob_severity_2[i],
          params$params$prob_severity_3[i],
          params$params$prob_severity_4[i])

      param_sample$disability_weight <- 
        c(params$params$disability_weight_1[i],
          params$params$disability_weight_2[i],
          params$params$disability_weight_3[i],
          params$params$disability_weight_4[i])     

      param_sample$incidence_rate <- params$params$incidence_rate[i]
      param_sample$case_fatality_ratio <- rep(c(params$params$case_fatality_ratio_1[i], params$params$case_fatality_ratio_2[i]), c(15, 86))
      
      res <- vaccine_impact_tf(disease = dis,
                               country = cntry,
                               population_data = population_data,
                               vaccine_coverage_data = vacc_cov_data,
                               vacc_scenario = vacc_scenario,
                               year = eval(parse(text = year)),
                               life_expectancy = life_exp,
                               parameter_sample = param_sample) 
                        
      cohort[[i]] <- res$cohort_size
      cases[[i]] <- res$cases
      deaths[[i]] <- res$deaths
      dalys[[i]] <- res$dalys 
    }
    
    stoch <- list()
    stoch$var <- c("cohort_size", "cases", "deaths", "dalys")
    stoch$val[[1]] <- cohort[[1]] # cohort size remains constant across nruns
    stoch$val[[2]] <- cases
    stoch$val[[3]] <- deaths
    stoch$val[[4]] <- dalys
    
    stoch_output[[k]] <- 
      vimc_stoch_report(disease = dis, country = cntry, sim_output = stoch)
    
    # stoch_parameters[[k]] <- stoch_param
    
    names(stoch$val) <- stoch$var
    central_output[[k]] <- vimc_central_report(disease = dis, country = cntry, 
                               sim_output = stoch$val)
  }
  saveRDS(stoch_output, paste0("outputs/stoch_output_list_", tstamp, "_", dis, "_", vacc_scenario_save, ".rds"))
   saveRDS(central_output, paste0("outputs/central_output_list_", tstamp, "_", dis, "_", vacc_scenario_save, ".rds"))
}
```


### Manual conversion2

```{r}
# data in which results from each country is stored in a list
devtools::load_all()

tstamp <- "20220514"
vacc <- "routine_2"
dis <- "Typhoid"
stocen <- "stoch" 
fls <- list.files("outputs", pattern = paste0(".*", stocen, ".*output.*list.*", tstamp, ".*", dis, ".*", vacc, ".*cid.*.*rds$"), full.names = T)
# fls <- fls[c(1:4, 10:11, 5:9)]
# res_list <- readRDS(fls[1])
# bind each element of the list by row

res <- lapply(fls, function(x) do.call(rbind, readRDS(x)))
res <- do.call(rbind, res)

# stocen <- "stoch"
# res <- readRDS(paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tstamp, ".rds"))

saveRDS(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tstamp, ".rds"))

fwrite(res, paste0("outputs/", stocen, "_", dis, "_", vacc, "_", tstamp, ".csv"))
  
```

### Auto-conversion of multiple files
```{r}
disease <- c("Typhoid")
vacc_scenario <- c("campaign_2")
tstamp <- "20211217T1"
library(data.table)
for(dis in disease) {
  for (vacc in vacc_scenario) {
    central <- list.files("outputs", pattern = paste0("central.*list.*", tstamp, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)
    ## bind by row from a list of data with the same format
    res <- do.call(rbind, readRDS(central))
    saveRDS(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".rds"))
    fwrite(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".csv"))
    
    stoch <- list.files("outputs", pattern = paste0("stoch.*output.*list.*", tstamp, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)
    res_stoch <- do.call(rbind, readRDS(stoch))
    saveRDS(res_stoch, paste0("outputs/stoch_", dis, "_", vacc, "_", tstamp, ".rds"))
    fwrite(res_stoch, paste0("outputs/stoch_", dis, "_", vacc, "_", tstamp, ".csv"))
    }
}
```
