---
title: "typhoid_practice_runs"
author: "Jong-Hoon Kim"
date: "9/14/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Vaccine Impact Modeling Consortium data
```{r}
devtools::load_all()

library(usethis)
# prob_dying_age <- fread('inst/extdata/202110gavi-1_dds-201910_p_dying_both.csv')
# use_data(prob_dying_age, overwrite = T)
# remove p_dying_age as prob_dying_age sounds clearer
# rm("data/p_dying_age.rda") this does not work, why? 
dis <- "typhoid"
vacc_cov_data <- fread('inst/extdata/coverage_202108test-1_typhoid-routine-default.csv')
# test_countries <- c("China", "Pakistan", "India", "Nigeria", "Ethiopia")
population_data <- as.data.frame(fread("outputs/population_wide.csv"))
# population_data$country <- clean_country_names(population_data$country)
# fwrite(population_data, "outputs/population_wide.csv")
parameter_data <- fread("inst/extdata/parameters.csv")

str_dis_severity <- c("Proportion of moderate typhoid fever",
"Proportion of severe typhoid fever",
"Proportion of severe typhoid fever with gastrointestinal bleeding",
"Proportion of typhoid fever with abdominal complications (other than gastrointestinal bleeding)")

str_dis_duration <- c("Duration of moderate typhoid fever",
"Duration of severe typhoid fever",
"Duration of severe typhoid fever with gastrointestinal bleeding",
"Duration of  typhoid fever with abdominal complications (other than gastrointestinal bleeding)")

str_dis_disability <- c("Disability weight for moderate typhoid fever",
"Disability weight for severe typhoid fever",
"Disability weight for severe typhoid fever with gastrointestinal bleeding",
"Disability weight for severe typhoid fever with abdominal complications (other than gastrointestinal bleeding)")

# # incidence data
# incid_data_typhoid <- fread("data/IHME-GBD_2019_DATA-3268a4a1-1.csv")
# incid_data_typhoid$country <- clean_country_names(incid_data_typhoid$location) 
# usethis::use_data(incid_data_typhoid, overwrite = T)

cntry <- unique(vacc_cov_data$country)

incid_data_typhoid %>% 
  filter.(measure == "Incidence", year == 2015, location %in% cntry) -> ir_data 

incid_data_typhoid %>% 
  filter.(measure == "Deaths", year == 2015, location %in% cntry) -> death_data 

incid_data_typhoid %>% 
  filter.(measure == "YLDs (Years Lived with Disability)",
          year == 2015, location %in% cntry) -> yld_data 

incid_data_typhoid %>% 
  filter.(measure == "YLLs (Years of Life Lost)",
          year == 2015, location %in% cntry) -> yll_data 

# age category <1 year, 1 to 4, 5 to 9, ..., 90 to 94  
# make age categories as in the incidence rate data (ir_data)
spread_age_IHME <- function(data, age_total = 101, variable = "val") {
  var <- match.arg(variable, c("val", "upper", "lower"))
  cat_first <- "<1 year"
  # next category starts "1 to 4", "5 to 9", ..., "90 to 94"
  age0 <- c(1, seq(5, 90, by = 5))
  age1 <- c(4, seq(9, 94, by = 5))
  age_cat <- c(paste0(age0, " to ", age1))
  
  new_data <- rep(NA, age_total)
  new_data[1] <- unlist(data[data$age == cat_first, var])
  for (ag in age0) {
    new_data[(ag+1):(ag+5)] <- 
      unlist(data[data$age == age_cat[which(age0 == ag)], var])
  } 
  # 95-100 age groups were set to be the same as "90 to 94"
  new_data[96:age_total] <- new_data[95]
  
  return(new_data)
}

extract_param_tf <- function(params_data, str, var_name = "value"){
  var <- match.arg(var_name, c("value", "upper_bound", "lower_bound"))
  params <- rep(NA, length(str))
  params_data %>% filter(tolower(disease) == tolower(dis)) -> dat
  for (i in 1:length(params)) {
    dat %>% 
    filter(definition == str[i]) %>%
    select(eval(var)) %>% 
    unlist() -> params[i]
  }
  return(params)
}

```


## Target countries and vaccine coverages
```{r setup, include=FALSE}
devtools::load_all()
# load additional packages
# library(tidyverse)

dis <- "typhoid"

# test_countries <- c("China", "Pakistan", "India", "Nigeria", "Ethiopia")
test_countries <- unique(vacc_cov_data$country)
cntry <- test_countries[1]

pop <- setup_cohorts(country = cntry)

vc <- create_vaccine_coverage(disease = dis,
                        country = cntry,
                        population = pop,
                        vaccine_coverage_data = vacc_cov_data)

vacc_protected <- calculate_vaccine_protected(disease = dis,
                                         country = cntry,
                                         population = pop,
                                         year = 2000:2100,
                                         vaccine_coverage = vc,
                                         vaccine_efficacy = 0.5,
                                         vaccine_immunity_duration = 15,
                                         exponential_decay = TRUE,
                                         vaccine_efficacy_by_year = NULL)

pop_risk <- pop - vacc_protected
sum(pop_risk < 0)


prop_high_risk  <- prop_basic_san
prop_high_risk[, prop := 1 - prop]


pars_data %>% 
  filter(tolower(disease) == dis) %>% 
  filter(definition == "relative risk of infection from at least basic sanition") %>%
  select(value) %>% 
  unlist() -> rr

rr <- 1 / rr # make the relative risk for the high-risk group

ir <- spread_age_IHME(ir_data)

cases <- calculate_cases(population = pop, 
                         country = cntry, 
                         disease = dis,
                         year_risk_ref = 2015,
                         prop_high_risk = prop_high_risk, 
                         incidence_rate = ir,
                         risk_ratio = rr)
pop[1:10,1]
cases[1:10,1]
head(pop)
head(cases)


dis_sev <- extract_param_tf(params_data = pars_data, str = str_dis_severity)
dis_dur <- extract_param_tf(params_data = pars_data, str = str_dis_duration)
dis_wt <- extract_param_tf(params_data = pars_data, str = str_dis_disability)

yld <- calculate_YLD_tf(cases = cases,
                        disability_weight = dis_wt,
                        prob_severity = dis_sev,
                        illness_duration = dis_dur)
yld[1:10,1:20]


ir_data$val[1:10]* sum(sev * wt * dur)
yld_data$val[1:10]

## age-specific incidence rate

deaths <- spread_age_IHME(death_data)
cfr <- deaths / ir

yll <- calculate_YLL_tf(cases = cases,
                        country = cntry,
                        case_fatality = cfr)


yll[1:10,"2015"]
cases[1:10, "2015"] * cfr[1:10] * life_expectancy[1:10, "2015"] # 0 -9 y0

ir[1:10] * cfr[1:10] * life_expectancy[1:10, "2015"] # 0 -9 y0
yll_data[yll_data$age == "1 to 4",]$val
yll_data[yll_data$age == "5 to 9",]$val
yll_data[yll_data$age == "<1 year",]$val
# it appears that life expectancy is lower for us compared to the IHME

daly <- calculate_DALY_tf(country = cntry,
                          cases = cases,
                          disability_weight = wt,
                          prob_severity = sev,
                          illness_duration = dur, 
                          case_fatality = cfr)
daly[1:10, 1:10]
yll[1:10, 1:10] + yld[1:10, 1:10]
```

## Population at risk
### 1. Use open defecation as the risk indicator
### 2. Use no access to improved water as the risk indicator
### 3. Use no access to improved sanitation as the risk indicator
```{r}

```

## Incidence rate and case fatality ratio
Incidence rate estimates from IHME were downloaded from the website: 
These country-level estimates were divided into two values of incidence in
population who have access to the risk factor, which we defined above, and those
who have not.  Then, since we have the estimates of the level of access to  
```{r}
ref_year <- 2015 # incidence rate in 2015 of IHME data
incid_data_typhoid %>% 
  filter.(measure == "Incidence", year == ref_year, country %in% cntry)  %>% 
  spread_age_IHME -> ir

```


## Sampling parameters

```{r}
# Sampling based on Sobol low discrepancy sequence   
test_countries
nruns <- 5
params_list <- set_params_tf(nruns = nruns,
                              country = test_countries,
                              parameter_data = pars_data,
                              incidence_rate_data = incid_data_typhoid)

# get country-level fixed parameters
life_exp <- 
  get_life_expectancy(life_expectancy_data = life_expectancy_data,
                      country = cntry, year = yr)

# fitting results on the proportion of population with access to basic sanitation
prop_high_risk  <- prop_basic_san
prop_high_risk[, prop := 1 - prop]
  
vi <- vaccine_impact_tf(disease = dis,
                  country = cntry,
                  population_data = population_data,
                  vaccine_coverage_data = vacc_cov_data,
                  vacc_scenario = "routine",
                  year = 2000:2100,
                  life_expectancy = life_exp,
                  prop_high_risk = prop_high_risk,
                  parameter_samples = params_list) 

```

## Simulation
Some methods require disease and other don't. How do I streamline it?

```{r}
# generate parameter sets to explore
# run vaccine impact
devtools::load_all()
set.seed(1234)
nruns <- 30
#VIMC report requires capital T in Typhoid
dis <- "Typhoid"
year <- "2000:2100"
vaccination_scenarios <- c("novacc", "routine")
tstamp <- format(Sys.time(), "%Y%m%d")
# target countries
template <- fread("inst/extdata/central-burden-template.202108test-1.Typhoid_IVI-Kim_standard.csv")
target_countries <- clean_country_names(unique(template$country_name))
# parameters to explore
parameter_data <- fread("inst/extdata/parameters.csv")

population_data <- as.data.frame(fread("outputs/population_wide.csv"))
library(magrittr)
parameter_data %<>% filter(tolower(disease) == tolower(dis))
# life expectancy data
life_expectancy_data <-
  fread("inst/extdata/202110gavi-1_dds-201910_2_life_ex_both.csv")
vaccine_coverage_data <- 
  fread("inst/extdata/coverage_202108test-1_typhoid-routine-default.csv")

# proportion of high-risk population data
prop_high_risk_data <- prop_basic_san #internal model fit results
prop_high_risk_data[, prop := 1 - prop]

str_dis_severity <- c("Proportion of moderate typhoid fever",
"Proportion of severe typhoid fever",
"Proportion of severe typhoid fever with gastrointestinal bleeding",
"Proportion of typhoid fever with abdominal complications (other than gastrointestinal bleeding)")

str_dis_duration <- c("Duration of moderate typhoid fever",
"Duration of severe typhoid fever",
"Duration of severe typhoid fever with gastrointestinal bleeding",
"Duration of  typhoid fever with abdominal complications (other than gastrointestinal bleeding)")

str_dis_disability <- c("Disability weight for moderate typhoid fever",
"Disability weight for severe typhoid fever",
"Disability weight for severe typhoid fever with gastrointestinal bleeding",
"Disability weight for severe typhoid fever with abdominal complications (other than gastrointestinal bleeding)")

# Sampling based on Sobol low discrepancy sequence  
params_list <- list()
for (k in 1:length(target_countries)) {
  cntry <- target_countries[[k]]
  params_list[[k]] <- 
    set_params_tf(nruns = nruns,
                  country = cntry,
                  parameter_data = parameter_data,
                  incidence_rate_data = incid_data_typhoid)
}
t
saveRDS(params_list, paste0("outputs/params_", dis, "_", tstamp, ".rds"))

stoch_output <- list()
stoch_parameters <- list()
central_output <- list()

# r <- 2
# k = 1
# i = 1
for (r in seq_along(vaccination_scenarios)) {
  vacc_scenario <- vaccination_scenarios[r]
  for (k in seq_along(target_countries)) {
    cntry <- target_countries[k]
    cohort <- list()
    cases <- list()
    deaths <- list()
    dalys <- list() 

    life_exp <- 
      get_life_expectancy(life_expectancy_data = life_expectancy_data,
                          country = cntry,
                          year = eval(parse(text = year)))
    prop_high_risk <- 
      prop_high_risk_data[prop_high_risk_data$country == cntry, ]
    
    message(paste0("vaccine scenario = ", vacc_scenario, ", country = ", cntry))
    ## country-specific parameters
    params <- params_list[[k]] # country by index
    
    for (i in 1:nruns) {
      # message(paste0("r = ", r, ", k = ", k, ", i = ", i))
      param_sample <- list()
      param_sample$vacc_efficacy <- params[[3]]$vacc_efficacy[i]
      param_sample$duration_vacc_protection <- 
        params[[3]]$duration_vacc_protection[i]
      param_sample$odd_ratio <- params[[3]]$odds_ratio[i]
      param_sample$illness_duration <- 
        c(params[[3]]$illness_duration_1[i],
          params[[3]]$illness_duration_2[i],
          params[[3]]$illness_duration_3[i],
          params[[3]]$illness_duration_4[i])

      param_sample$prob_severity <- 
        c(params[[3]]$prob_severity_1[i],
          params[[3]]$prob_severity_2[i],
          params[[3]]$prob_severity_3[i],
          params[[3]]$prob_severity_4[i])

      param_sample$disability_weight <- 
        c(params[[3]]$disability_weight_1[i],
          params[[3]]$disability_weight_2[i],
          params[[3]]$disability_weight_3[i],
          params[[3]]$disability_weight_4[i])     

      param_sample$incidence_rate <- params[[1]][, i]
      param_sample$case_fatality_ratio <- params[[2]][, i]
      
      res <- vaccine_impact_tf(disease = dis,
                               country = cntry,
                               population_data = population_data,
                               vaccine_coverage_data = vaccine_coverage_data,
                               vacc_scenario = vacc_scenario,
                               year = 2000:2100,
                               life_expectancy = life_exp,
                               prop_high_risk = prop_high_risk,
                               parameter_sample = param_sample) 
                        
    
      cohort[[i]] <- res$cohort_size
      cases[[i]] <- res$cases
      deaths[[i]] <- res$deaths
      dalys[[i]] <- res$dalys 
    }
    
    stoch <- list()
    stoch$var <- c("cohort_size", "cases", "deaths", "dalys")
    stoch$val[[1]] <- cohort[[1]] # cohort size remains constant across nruns
    stoch$val[[2]] <- cases
    stoch$val[[3]] <- deaths
    stoch$val[[4]] <- dalys
    
    stoch_output[[k]] <- 
      vimc_stoch_report(disease = dis, country = cntry, sim_output = stoch)
    
    # stoch_parameters[[k]] <- stoch_param
    
    names(stoch$val) <- stoch$var
    central_output[[k]] <- vimc_central_report(disease = dis, country = cntry, 
                               sim_output = stoch$val)
  }
  saveRDS(stoch_output, paste0("outputs/stoch_output_list_", tstamp, "_", dis, "_", vacc_scenario, ".rds"))
   saveRDS(central_output, paste0("outputs/central_output_list_", tstamp, "_", dis, "_", vacc_scenario, ".rds"))
}
```


## Combine results into a single file
Results across countries given vaccination scenario (i.e., novacc, campaign, or
routine) are in a list type. The do.call(rbind, x) is applied to make them a 
single data.frame   
```{r}
disease <- c("typhoid")
vacc_scenario <- c("novacc", "routine")
tstamp <- "20210930"
library(data.table)
for(dis in disease) {
  for (vacc in vacc_scenario) {
    central <- list.files("outputs", pattern = paste0("central.*list.*", tstamp, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)
    res <- do.call(rbind, readRDS(central))
    saveRDS(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".rds"))
    fwrite(res, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".csv"))
    
    stoch_output <- list.files("outputs", pattern = paste0("stoch.*output.*list.*", tstamp, ".*", dis, ".*", vacc, ".*rds$"), full.names = T)
    res_stoch <- do.call(rbind, readRDS(stoch_output))
    saveRDS(res_stoch, paste0("outputs/stoch_output_", dis, "_", vacc, "_", tstamp, ".rds"))
    fwrite(res_stoch, paste0("outputs/stoch_output_", dis, "_", vacc, "_", tstamp, ".csv"))
    }
}
```

## Diagnostic plots
```{r}
library(ggplot2)
tstamp <- "20210930"
for (cntry in target_countries) {
  dis <- "typhoid"
  vacc <- "routine"
  # vacc <- "novacc"
  # cntry <- "Iran, Islamic Republic of"
  res <- readRDS(paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".rds"))
  plt <- year_val_plot(disease = dis, country = cntry, sim_central_value = res)
  # year_val_plot(disease = dis, country = cntry, sim_central_value = res, age = 10)
  print(plt)
  invisible(readline(prompt = "Press RET to continue"))
} 
```

## Diagnostic plots II
Visualize cohort size, cases, deaths, and dalys across vaccination scenarios for each of the countries over the period of 2000 - 2100
```{r}
library(ggplot2)
library(grid)
library(gridExtra)
dis <- "typhoid"
tstamp <- "20210930"
for (cntry in target_countries){
  res1 <- readRDS(paste0("outputs/central_",  dis, "_", "novacc_", tstamp, ".rds"))
  res2 <- readRDS(paste0("outputs/central_",  dis, "_", "routine_", tstamp, ".rds"))
  
  p1 <- year_val_plot(disease = dis, country = cntry, sim_central_value = res1, vacc_scenario = "novacc")
  p2 <- year_val_plot(disease = dis, country = cntry, sim_central_value = res2, vacc_scenario = "routine")

  plt <- grid.arrange(p1, p2, ncol = 1)
  print(plt)
  invisible(readline(prompt = "Press RET to continue"))
} 
```

## Diagnostic III
1. Check if there are population categories (defined by chronological year and age) have negative numbers
2. Check if cohort sizes are same across vaccination scenarios 
3. Check if cases, deaths, and dalys are larger for non-vaccination scenarios compared to vaccination scenarios
3. Check if cases, deaths, and dalys are larger for routine (or campaign) scenario compared to the 
```{r}
dis <- "typhoid"
tstamp <- "20210930"
for (cntry in target_countries){
  res1 <- readRDS(paste0("outputs/central_",  dis, "_", "novacc_", tstamp, ".rds"))
  res2 <- readRDS(paste0("outputs/central_",  dis, "_", "routine_", tstamp, ".rds"))
  
  if (sum(res1$cohort_size != res2$cohort_size) != 0) {
    cat("Cohort size of novacc and routine scenrios are different for", cntry, "\n")
  }
  if (sum(res1$cases < res2$cases) != 0) {
    cat("Cases are bigger for routine scenario than novacc scenrios for", cntry, "\n")    
  }
   if (sum(res1$deaths < res2$deaths) != 0) {
    cat("Deaths are bigger for routine scenario than novacc scenrios for", cntry, "\n")    
  }
  if (sum(res1$dalys < res2$dalys) != 0) {
    cat("DALYs are bigger for routine scenario than novacc scenrios for", cntry, "\n")
  }
}
```

## Prepare csv files for uploading the results onto Montagu
```{r}
library(data.table)
dis <- "Typhoid"
tstamp <- "20210930"

for (vacc in c("novacc", "routine")) {
  fl <- list.files("outputs/", pattern = paste0("central.*", dis, ".*", vacc, ".*", tstamp,  ".*rds$"), full.names = T)
  d <- readRDS(fl)
  fwrite(d, paste0("outputs/central_", dis, "_", vacc, "_", tstamp, ".csv"))
}

# fls <- list.files("outputs", pattern = paste0("^param_combined.*", tstamp, ".*", dis, ".*", ".*rds$"), full.names = T)
# x <- lapply(fls, function(x) readRDS(x))
# d <- do.call(rbind, x)
# fwrite(d, paste0("outputs/param_", dis, "_combined", tstamp, ".csv"))
```


## Combine parameters
```{r}
lst <- list()
for (i in 1:length(target_countries)) {
  params <- params_list[[i]]
  
  ir <- params[[1]]
  t_ir <- data.table::transpose(ir)
  colnames(t_ir) <- paste0("incid_rate_ag_", 0:100)
  cfr <- params[[2]]
  t_cfr <- data.table::transpose(cfr)
  colnames(t_cfr) <- paste0("case_fatality_ratio_ag_", 0:100)
  
  lst[[i]] <- cbind(params[[3]], t_ir, t_cfr)
}
res <- do.call(rbind, lst)
saveRDS(res, paste0("outputs/params_", dis, "_", tstamp, ".rds"))
fwrite(res, paste0("outputs/params_", dis, "_", tstamp, ".csv"))
```

# Diagnostic IV: Negative population size
Negative values can happen when there is population size for a certain age group
is zero and in the previous year there are vaccinated people in one-year younger
age group some value and model predicts some people move to the next age group,
which is determined by the probability of dying and aging, vaccine  

```{r}
vacc_scenario <- c("novacc", "routine")
tstamp <- "20210930"

for (vacc in vacc_scenario) {
  fl <- paste0("outputs/central_typhoid_", vacc, "_", tstamp, ".csv")
  message(paste0("file = ", fl))
  d <- fread(fl)
  if(sum(d$cohort_size < 0) > 0){
    message("Negative cohort size!")
  }
  if(sum(d$cases < 0) > 0){
    message("Negative cases!")
  }
  if(sum(d$deaths < 0) > 0){
    message("Negative deaths!")
  }
  if(sum(d$dalys < 0) > 0){
    message("Negative dalys!")
  }
}
```
